---
title: "2016 release"
output: html_document
---
install.packages("curl")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("gplots")
install.packages("corrplot")
install.packages("ggplot2")
install.packages("vegan")
install.packages("data.table")
install.packages("tidyr")
install.packages("formattable")

```{r}
save.image(file='myEnvironment.RData')
dir()

library(curl)
library(dplyr)
library(tidyverse)
library(gplots)
library(corrplot)
library(ggplot2)
library(vegan)
library(data.table)
library(tidyr)
library(formattable)
```

## Importing the pre-focals datasheet (NOT FULL)
```{r}
pre_focals<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/2016_pre-release_FOCALS.csv") #this one only has months 6-9!! NOT full
pre_focals<- read.csv(pre_focals, header = T, na.strings=c(""," ","NA"))
pre_focals <- pre_focals %>% mutate_all(na_if,"")
levels(pre_focals$BEHAVIOUR.2)
head(pre_focals)
nrow(pre_focals)
summary(pre_focals)
names(pre_focals)
unique(pre_focals$FOCAL.ID)
n_distinct(pre_focals$FOCAL.ID)
unique(pre_focals$AGE)
sort(table(pre_focals$BEHAVIOUR), decreasing = TRUE)
n_distinct(pre_focals$FOCAL.REFERENCE.CODE) #1012 total focals
```

## Importing the FULL pre-focals datasheet
```{r}
pre_focals<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/2016_pre-release_focals_FULL.csv")
pre_focals<- read.csv(pre_focals, header=T, na.strings=c(""," ","NA"))
levels(pre_focals$BEHAVIOUR)[levels(pre_focals$BEHAVIOUR) == "F"] <- "FE" #changing the Fs to FEs
unique(pre_focals$BEHAVIOUR)

pre_focals_observers<- pre_focals %>%
  gather("X","Observers", 13:14, na.rm= T)
unique(pre_focals_observers$Observers)
```

## Importing the pre-continuous dataset (this is NOT FULL)
```{r}
pre_cont<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/2016_pre-release_CONT.csv")
pre_cont<- read.csv(pre_cont, header=T, na.strings=c(""," ","NA"))
pre_cont$X<- NULL #There was an empty column in front with all NAs
pre_cont<- pre_cont[-c(1),] #the first row was just an empty row of NAs
head(pre_cont)
```

## Importing the FULL pre-continuous dataset
```{r}
pre_cont<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/2016_pre-release_cont_FULL.csv")
pre_cont<- read.csv(pre_cont, header=T, na.strings=c(""," ","NA"))
unique(pre_cont$MONTH)
```

# Importing the post-focals dataset (Not full?)
```{r}
post_focals<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/2016_post-release_FOCALS.csv")
post_focals<- read.csv(post_focals, header= T, na.strings=c(""," ","NA"))
head(post_focals)
```

## Creating an Age/Sex plot
```{r}
age_sex_plot <- ggplot(pre_focals, aes(sex, group = age)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~age)
age_sex_plot #There are more female than male adults, almost all subadults are males, and more juveniles are females... overall more females than males 
```

## POPs
## Filtering out Pops from the pre-focals dataset
```{r}
Pops<- pre_focals %>% filter(FOCAL.ID == "PO")
head(Pops)
n_distinct(Pops$MONTH)
#unique(Pops$FOCAL.REFERENCE.CODE)
```

### Testing to see if I get the same SD value if I 1) do three focals separately and then find the average SD and 2)combine all three focals into one df and then find the SD for that 
```{r}
#These three chunks are each focal separately
Pops_behaviorcount<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10101") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount

Pops_behaviorcount2<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10407") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount2

Pops_behaviorcount3<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "20109") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount3

#Finding SD for each focal
SD1<- diversity(Pops_behaviorcount$count)
SD2<- diversity(Pops_behaviorcount2$count)
SD3<- diversity(Pops_behaviorcount3$count)

SD1
SD2
SD3
SDs<- c(SD1,SD2,SD3) #combining the focals
AvgSD<- mean(SDs)
AvgSD #1.145476

#Now I'm filtering out all 3 focals so that it counts the behaviors across all three
Pops_behaviorcount4<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE %in% c("10101","10407","20109")) %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount4

#finding the SD for this one df
SD4<- diversity(Pops_behaviorcount4$count)
SD4 #1.307044

#The SD values are DIFFERENT for the two methods! 
```

### Getting one SD value for all of Pops 56 focals... I don't think this is a reliable value?
```{r}
Pops_behaviorcountTEST2<- Pops %>% 
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcountTEST2

DIVTEST<- diversity(Pops_behaviorcountTEST2$count)
DIVTEST #1.459
```

### Playing around with the two behavior columns for one of Pop's focals, I need to somehow incorporate the behaviors from the Behavior.2 column into my SD calculations (see code below for proper code)
```{r}
Pops_behaviorcount2<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10407") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount2

Pops_behaviorcount3<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10407") %>%
  drop_na(BEHAVIOUR.2) %>%
  group_by(BEHAVIOUR.2) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount3

#!is.na(BEHAVIOUR.2)
#na.rm = TRUE
```

## THIS IS THE PROPER CODE, just practicing it out on individual focals first
```{r}
Pops_behaviorcount4<- Pops %>% 
  filter(focal.referenece == "10407") %>%
  group_by(BEHAVIOUR, BEHAVIOUR.2)
  summarize(count=n())
Pops_behaviorcount4

Pops_behaviorcount4 <- Pops_behaviorcount4 %>%
  gather("Obs", "behavior", 19:20, na.rm = T)
  
Pops_behaviorcount5<- Pops_behaviorcount4 %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount5
```


### Saving the image as a jpeg
```{r}
# pdf(file = "/New User/Desktop/SD Plot NEW.pdf",   # The directory you want to save the file in
    #width = 4, # The width of the plot in inches
    #height = 4) # The height of the plot in inches

jpeg("SDplot NEW.jpg", width = 900, height = 500)
x<- barplot(pre_SDs_table, beside = TRUE, space = c(0.25,1), xlab = "", ylab = "Shannon Diversity Index", names.arg = c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman"), ylim = c(0,2), cex.lab = 1.25, font.lab= 2, cex.names = 1, las=2, col="#339999")#I don't love this barplot visual, I can't figure out how to turn the x-axis labels on an angle using a table
x

dev.off()
```

## Pre-Release Vigilance Counts from each Individual
```{r}
Pops_vig<- Pops %>% 
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% #filtering out only the vigilance behavior from this new "behavior" column created by gather()
  summarize(count=n()) #counting all the vigilance entries
Pops_vig$prop<- Pops_vig$count/nrow(Pops)
Pops_vig$ID<- "Pops"
Pops_vig

Zip_vig<- Zip %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Zip_vig$prop<- Zip_vig$count/nrow(Zip)
Zip_vig$ID<- "Zip"

Jack_vig<- Jack %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Jack_vig$prop<- Jack_vig$count/nrow(Jack)
Jack_vig$ID<- "Jack"

Blue_vig<- Blue %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Blue_vig$prop<- Blue_vig$count/nrow(Blue)
Blue_vig$ID<- "Blue"

Alex_vig<- Alex %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Alex_vig$prop<- Alex_vig$count/nrow(Alex)
Alex_vig$ID<- "Alex"

Bart_vig<- Bart %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Bart_vig$prop<- Bart_vig$count/nrow(Bart)
Bart_vig$ID<- "Bart"

Nev_vig<- Nev %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Nev_vig$prop<- Nev_vig$count/nrow(Nev)
Nev_vig$ID<- "Neville"

Aug_vig<- Aug %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Aug_vig$prop<- Aug_vig$count/nrow(Aug)
Aug_vig$ID<- "Aug"

Amy_vig<- Amy %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Amy_vig$prop<- Amy_vig$count/nrow(Amy)
Amy_vig$ID<- "Amy"

May_vig<- May %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
May_vig$prop<- May_vig$count/nrow(May)
May_vig$ID<- "May"

Toni_vig<- Toni %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Toni_vig$prop<- Toni_vig$count/nrow(Toni)
Toni_vig$ID<- "Toni"

Boo_vig<- Boo %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Boo_vig$prop<- Boo_vig$count/nrow(Boo)
Boo_vig$ID<- "Boo"

Bgm_vig<- Bgm %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Bgm_vig$prop<- Bgm_vig$count/nrow(Bgm)
Bgm_vig$ID<- "Big Mama"

Kovu_vig<- Kovu %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Kovu_vig$prop<- Kovu_vig$count/nrow(Kovu)
Kovu_vig$ID<- "Kovu"

Ed_vig<- Ed %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Ed_vig$prop<- Ed_vig$count/nrow(Ed)
Ed_vig$ID<- "Eddy"

Tink_vig<- Tink %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Tink_vig$prop<- Tink_vig$count/nrow(Tink)
Tink_vig$ID<- "Tinker"

Bat_vig<- Bat %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Bat_vig$prop<- Bat_vig$count/nrow(Bat)
Bat_vig$ID<- "Batman"

Mango_vig<- Mango %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
Mango_vig$prop<- Mango_vig$count/nrow(Mango)
Mango_vig$ID<- "Mango"
```

## Merging Pre-release vigilance for all individuals
```{r}
vig_df_ALL<- list(Pops_vig, Zip_vig, Jack_vig, Blue_vig, Alex_vig, Bart_vig, Nev_vig, Aug_vig, Amy_vig, May_vig, Toni_vig, Boo_vig, Bgm_vig, Kovu_vig, Ed_vig, Tink_vig, Bat_vig, Mango_vig) 
vig_df_ALL<- Reduce(function(x, y) merge(x, y, all=TRUE), vig_df_ALL, accumulate=FALSE) #Not sure how this works but this is needed to make it into a nice df

#This method works mostly (minus the legend) but I'm going to try with ggplot below
jpeg("Vig Plot.jpg", width = 900, height = 500)
vig_plot<- barplot(vig_df_ALL$prop, names.arg = vig_df_ALL$ID, cex.lab = 1, font.lab= 2, cex.names = 0.8, las=2, col="#339999", ylab = "Vigilance Frequency", ylim= c(0,1), density=c(100,100,100,100,100,100,100,100,100,100,20,100,20,20,100,20,20,20), legend = TRUE, args.legend = list(bty="n", x="topleft")) #Can't get the legend to work here!
vig_plot

# This plot looks good!!
c4 = c("Female", "Male", )
vig_df_ALL = cbind(vig_df_ALL, c4)

#adding in a new column for sex to try to fill the bar colors by it
vig_df_ALL<- vig_df_ALL %>% 
  add_column(sex = c("F","F","F","F","F","F","F","F","M","F","M","F","M","M","F","M","M","M"), .after="ID")

#The fill in line one doesn't work
ggplot(vig_df_ALL, aes(x = reorder(ID, -prop), y = prop), fill= sex) + #the reorder function with -prop reorders the x-axis to go from highest to lowest
  geom_bar(stat = "identity") + 
  scale_y_continuous(labels=scales::percent) + #this changes the y-axis to percents
  theme(axis.text.x = element_text(angle = 45, hjust =1)) + #turns axis labels to 45 degrees and hjust get rid of excess space
  ylab("Percent Vigilance Behavior") +
  xlab("Individual") +
  scale_fill_manual("legend", values = c("F" = "red", "M" = "blue")) #doesn't want to work

# This also doesn't work
ggplot(vig_df_ALL, aes(x = ID, y = prop), fill= sex) + 
  geom_bar(stat = "identity") + 
  scale_y_continuous(labels=scales::percent) + #this changes the y-axis to percents
  theme(axis.text.x = element_text(angle = 45, hjust =1)) + #turns axis labels to 45 degrees and hjust get rid of excess space
  ylab("Percent Vigilance Behavior") +
  xlab("Individual") +
  scale_fill_manual("legend", values = c("F" = "red", "M" = "blue"))

dev.off()
```

*Pre-release Vigilance across entire troop*
```{r}
pre_focals_vig<- pre_focals %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>% 
  filter(behavior == "V") %>% 
  summarize(count=n())
pre_focals_vig$prop<- pre_focals_vig$count/nrow(pre_focals)
```

### Vigilance Dataframe Editing
```{r}
#first filtering out Batman since he's a wild male
pre_focals_edited<- pre_focals %>%
  filter(FOCAL.ID != "BT")

#creating a "vigilance-only" dataframe
pre_focals_vig<- pre_focals_edited %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>% #combining behavior1&2
  group_by(behavior) %>% 
  filter(behavior == "V") #pulling out vigilance
head(pre_focals_vig)
```

### Rate of Vigilance for ADULT FEMALES
```{r}
#1. Create a new dataframe with adult females only
pre_focals_Afems<- pre_focals_edited %>%
  filter(sex == "F", age == "A")

#2. create a character list of all adult female IDs
MonkeyIDs_preAfems<- as.character(unique(pre_focals_Afems$FOCAL.ID))
MonkeyIDs_preAfems

# 2. Get a list of dataframes (from the updated vigilance dataframe), subsetted by adult female monkey ID
monkey.list_preAfems<-lapply(MonkeyIDs_preAfems, function(x){pre_focals_vig[pre_focals_vig[["FOCAL.ID"]] == x, ]})
head(monkey.list_preAfems)

#3. Create new list of tibbles to look at vigilance counts per hour 
monkey.list_pre_vigAfems<-
  monkey.list_preAfems %>%
  purrr::map(~group_by(.,day,hour)) %>% #grouping by day&hour so that I get a vigilance count for each unique day-hour combination per individual 
  purrr::map(~summarize(.,count=n())) #counting up vigilance 
head(monkey.list_pre_vigAfems)

#4. Combining all the tibbles into one
monkey.list_pre_vigAfems<- monkey.list_pre_vigAfems %>% 
  purrr::reduce(dplyr::full_join) %>% #using full_join since all columns are the same across tibbles
  mutate(release = "pre") #creating a new column called "release"
monkey.list_pre_vigAfems
mean(monkey.list_pre_vigAfems$count) #10.35
```

### Rate of Vigilance for Subadults and Juveniles
```{r}
#1. Create a new dataframe with subadults/juvs only
pre_focals_young<- pre_focals_edited %>%
  filter(age %in% c("SA","J"))

#2. create a character list of all juvs/subs
MonkeyIDs_preyoung<- as.character(unique(pre_focals_young$FOCAL.ID))
MonkeyIDs_preyoung

# 2. Get a list of dataframes (from the updated vigilance dataframe), subsetted by sub/juv monkey ID
monkey.list_preyoung<-lapply(MonkeyIDs_preyoung, function(x){pre_focals_vig[pre_focals_vig[["FOCAL.ID"]] == x, ]})
head(monkey.list_preyoung)

#3. Create new list of tibbles to look at vigilance counts per hour 
monkey.list_pre_vigyoung<-
  monkey.list_preyoung %>%
  purrr::map(~group_by(.,day,hour)) %>% #grouping by day&hour so that I get a vigilance count for each unique day-hour combination per individual 
  purrr::map(~summarize(.,count=n())) #counting up vigilance 
head(monkey.list_pre_vigyoung)

#4. Combining all the tibbles into one
monkey.list_pre_vigyoung<- monkey.list_pre_vigyoung %>% 
  purrr::reduce(dplyr::full_join) %>% #using full_join since all columns are the same across tibbles
  mutate(release = "pre") #creating a new column called "release"
monkey.list_pre_vigyoung
mean(monkey.list_pre_vigyoung$count) #10.18
```

### Rate of Vigilance for Pops
```{r}
pre_focals_vigPops<- pre_focals_vig %>%
  filter(FOCAL.ID == "PO") %>%
  group_by(day,hour) %>%
  summarize(count=n()) %>%
  mutate(release = "pre")
  
mean(pre_focals_vigPops$count) #14.30
```


## Creating a DF for the pre-release Shannon Diversity Index values so that I can use the barplot() function more easily
```{r}
Individual<- c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman") #my first column
SD<- c(Pop_pre_avgSDs, Zip_pre_avgSDs, Jack_pre_avgSDs, Blue_pre_avgSDs, Alex_pre_avgSDs, Bart_pre_avgSDs, Nev_pre_avgSDs, Aug_pre_avgSDs, Amy_pre_avgSDs, May_pre_avgSDs, Toni_pre_avgSDs, Boo_pre_avgSDs, Bgm_pre_avgSDs, Kovu_pre_avgSDs, Ed_pre_avgSDs, Tink_pre_avgSDs, Bat_pre_avgSDs) #My second column
BD_df<- data.frame(Individual,SD) #Turn into data frame
BD_df

row.names(BD_df)<- c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman") #I need to customize my rownames because with the code I use down below to turn the names on an angle I need to use the paste(rownames) function... otherwise they would just be 1,2,3...
row.names(BD_df)

par(mar = c(11, 7, 2, 2) + 0.2) #add room for the rotated labels

end_point = 0.5 + nrow(BD_df) + nrow(BD_df) - 1 #this is the line which does the trick (together with barplot "space = 1" parameter)

barplot(BD_df$SD, col = "#69b3a2", 
        main = "",
        ylab = "Shannon Diversity Index", ylim = c(0,2.5),
        xlab = "",
        space = 0.75)
text(seq(1.5, end_point, by = 2), par("usr")[3]-0.25, 
     srt = 60, adj = 1, xpd = TRUE, #rotate 60 degrees (srt = 60)
     labels = paste(rownames(BD_df)), cex = 0.75)
```

### Creating a table by hand of the confirmed dead/presumed dead/presumed emigrated for males/females/infants
```{r}
info<- matrix(c(7,1,0,4,11,3,3,0,3,0,2,0,21,4,5,4), ncol=4, byrow = T)
rownames(info)<- c("Male", "Female", "Infant", "Totals")
colnames(info)<- c("Released", "Confirmed Dead", "Presumed Dead", "Presumed Emigrated")
info<- as.table(info)
info
```

### Using formattable() function to create a nicer table
```{r}
summary<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/Summary%20Table.csv")
summary<- read.csv(summary, header = T) #Had to create an excel table and import as df to use formattable()

formattable(summary)

customGreen0 = "#DeF7E9"

customGreen = "#71CA97"

customRed = "#ff7f7f"

formattable(summary, 
            align =c("l","c","c","c","c"), 
            list(`Vervets` = formatter(
              "span", style = ~ style(color = "grey",font.weight = "bold"))
))
```

# POPs
## Filtering out Pops from the post-focals dataset
```{r}
Pops2<- post_focals %>% filter(FOCAL.ID == "PO")
head(Pops2)
n_distinct(Pops2$FOCAL.REFERENCE.CODE) #64 distinct focals
```

## Looking at Pop's behaviors in month 3 post-release (March) 
```{r}
#Month 3
Pops2_behaviorcount_month3<- Pops2 %>% 
  filter(MONTH == "3") %>% #filtering out month "3"
  gather("Obs", "behavior", 19:20, na.rm = T) %>% #using the gather() function to create two new columns "obs" and "behavior" using columns 19 & 20 ("Behavior" & "Behavior.2"). This adjusts the rows and basically combines all of the behaviors into one column under "behavior"
  group_by(behavior) %>% #Now grouping by this new "behavior" column created by gather()
  summarize(count=n()) %>% #counting the different behaviors and arranging them in descending order
  arrange(desc(count))
Pops2_behaviorcount_month3
```

## Filtering out each individual from the pre-CONTINUOUS dataset
```{r}
Pops_c<- pre_cont %>% filter(FOCAL.ID == "PO")
Zip_c<- pre_cont %>% filter(FOCAL.ID == "ZI")
Mango_c<- pre_cont %>% filter(FOCAL.ID == "MG")
Jack_c<- pre_cont %>% filter(FOCAL.ID == "JA")
Bart_c<- pre_cont %>% filter(FOCAL.ID == "BA")
Nev_c<- pre_cont %>% filter(FOCAL.ID == "NE")
Blue_c<- pre_cont %>% filter(FOCAL.ID == "BL")
Alex_c<- pre_cont %>% filter(FOCAL.ID == "AL")
Aug_c<- pre_cont %>% filter(FOCAL.ID == "AU")
Amy_c<- pre_cont %>% filter(FOCAL.ID == "AM")
May_c<- pre_cont %>% filter(FOCAL.ID == "MA")
Toni_c<- pre_cont %>% filter(FOCAL.ID == "TO")
Boo_c<- pre_cont %>% filter(FOCAL.ID == "BO")
Bgm_c<- pre_cont %>% filter(FOCAL.ID == "BM")
Kovu_c<- pre_cont %>% filter(FOCAL.ID == "KO")
Eddy_c<- pre_cont %>% filter(FOCAL.ID == "ED")
Tink_c<- pre_cont %>% filter(FOCAL.ID == "TI")
Bat_c<- pre_cont %>% filter(FOCAL.ID == "BT")
```

## Activity Budgets
### Behavior Categories
```{r}
pre_focals_budget<- pre_focals %>%
  #filter(FOCAL.ID != "BT") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>% #gathering the two behavior columns
  filter(!behavior %in% c("AL","BL","PO")) %>% #removing these from the behavior column becuase they're mistakes
  mutate(behavior.category = #creating new column for behavior categories
           case_when(behavior %in% c("G+","G-","PR-","PR+","C","CL","N","NU","SU","PL","MA","MO") ~ as.character("Social"),
                     behavior %in% c("F","FE","FO") ~ as.character("Feeding"),
                     behavior %in% c("L","R") ~ as.character("Activity"),
                     behavior %in% c("V","PA") ~ as.character("Predator"),
                     behavior %in% c("A+","A-","TH+","TH-","MP+","MP-") ~ as.character("Dominance"),
                     behavior %in% c("YA","SC","SG","SM","PC") ~ as.character("Stress"),
                     behavior %in% c("PH","AH")~ as.character("Human"),
                     behavior %in% c("O","OS")~ as.character("Other")))
unique(pre_focals_budget$behavior.category)
pre_focals_budget %>% filter(behavior %in% c("PH","AH"))
```

### Mean Activity Budget Per Focal
```{r}
#Pops
Pops_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "PO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Pops_budget<- Pops_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Pops_budget$focal.reference))))
Pops_budget$total<- NULL

#Blue
Blue_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BL") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Blue_budget<- Blue_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Blue_budget$focal.reference))))
Blue_budget$total<- NULL

#Amy
Amy_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "AM") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Amy_budget<- Amy_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Amy_budget$focal.reference))))
Amy_budget$total<- NULL

#Alex
Alex_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "AL") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Alex_budget<- Alex_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Alex_budget$focal.reference))))
Alex_budget$total<- NULL

#Boo
Boo_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Boo_budget<- Boo_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Boo_budget$focal.reference))))
Boo_budget$total<- NULL

#Big Mama
Bgm_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BM") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Bgm_budget<- Bgm_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Bgm_budget$focal.reference))))
Bgm_budget$total<- NULL

#Augustine
Aug_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "AU") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Aug_budget<- Aug_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Aug_budget$focal.reference))))
Aug_budget$total<- NULL

#Bart
Bart_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Bart_budget<- Bart_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Bart_budget$focal.reference))))
Bart_budget$total<- NULL

#Eddy
Ed_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "ED") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Ed_budget<- Ed_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Ed_budget$focal.reference))))
Ed_budget$total<- NULL

#Jack
Jack_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "JA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Jack_budget<- Jack_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Jack_budget$focal.reference))))
Jack_budget$total<- NULL

#Tinker
Tink_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "TI") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Tink_budget<- Tink_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Tink_budget$focal.reference))))
Tink_budget$total<- NULL

#Kovu
Kovu_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "KO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Kovu_budget<- Kovu_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Kovu_budget$focal.reference))))
Kovu_budget$total<- NULL

#Mango
Mango_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "KO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Mango_budget<- Mango_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Mango_budget$focal.reference))))
Mango_budget$total<- NULL

#Neville
Nev_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "NE") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Nev_budget<- Nev_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Nev_budget$focal.reference))))
Nev_budget$total<- NULL

#Toni
Toni_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "TO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Toni_budget<- Toni_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Toni_budget$focal.reference))))
Toni_budget$total<- NULL

#May
May_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "MA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

May_budget<- May_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(May_budget$focal.reference))))
May_budget$total<- NULL

#Zip
Zip_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "ZI") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Zip_budget<- Zip_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Zip_budget$focal.reference))))
Zip_budget$total<- NULL
```

*Troop Mean Activity Budgets*
```{r}
pre_fullbudget<- Pops_budget %>%
  full_join(Blue_budget,by= "behavior.category") %>%
  full_join(Amy_budget,by= "behavior.category") %>%
  full_join(Alex_budget,by= "behavior.category") %>%
  full_join(Boo_budget,by= "behavior.category") %>%
  full_join(Bgm_budget,by= "behavior.category") %>%
  full_join(Aug_budget,by= "behavior.category") %>%
  full_join(Bart_budget,by= "behavior.category") %>%
  full_join(Ed_budget,by= "behavior.category") %>%
  full_join(Jack_budget,by= "behavior.category") %>%
  full_join(Tink_budget,by= "behavior.category") %>%
  full_join(Kovu_budget,by= "behavior.category") %>%
  full_join(Mango_budget,by= "behavior.category") %>%
  full_join(May_budget,by= "behavior.category") %>%
  full_join(Toni_budget,by= "behavior.category") %>%
  full_join(Nev_budget,by= "behavior.category") %>%
  full_join(Zip_budget,by= "behavior.category") %>%
  rowwise() %>% #this allows us to work with row-wise dataframe where we want to look at each row individually
  mutate(TroopMean = ((sum(c_across(2:18),na.rm=T))/17), #adding up all the means per category and dividing by 17 monkeys to get average
        release = "pre") #adding column to identify pre-release

pre_fullbudget<- pre_fullbudget[,c("behavior.category","TroopMean","release")] #these are the only columns I want to look at now
pre_fullbudget
```

### Activity Budget Bar Chart
```{r}
budget_plot<-ggplot(data=pre_fullbudget, aes(x=behavior.category, y=TroopMean)) +
  geom_bar(stat="identity")
budget_plot
```

### Batman vs Rehabilitant Males: Activity Budget 
```{r}
#Batman
Bat_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BT") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())
Bat_budget

Bat_budget<- Bat_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Bat_budget$focal.reference))))
Bat_budget$total<- NULL

Pops_budget_pred<- Pops_budget %>%
  filter(behavior.category == "Predator") %>%
  mutate(ID = "Pops")
Pops_budget_pred

Bat_budget_pred<- Bat_budget %>%
  filter(behavior.category == "Predator") %>%
  mutate(ID = "Batman")

Males_budget<- bind_rows(Pops_budget_pred, Bat_budget_pred)
Males_budget %>% filter(ID == "Pops")
```

### Trying to do Wilcoxon Test with Batman and Pops
```{r}
ggdensity(Males_budget$count)
ggqqplot(Males_budget$count)

ggboxplot(Males_budget, x = "ID", y = "count", 
          color = "ID", palette = c("#00AFBB", "#E7B800"),
          ylab = "count", xlab = "ID")

WT_PopsBat_pred<- wilcox.test(count ~ ID, data = Males_budget, paired = FALSE)
WT_PopsBat_pred
```

## Average Activity Budget for entire troop
```{r}
#2. create a character list of all individuals in the budget dataframe
MonkeyIDs_budget<- as.character(unique(pre_focals_budget$FOCAL.ID))
MonkeyIDs_budget

# 2. Get a list of dataframes (from the budget dataframe), subsetted by individual
monkey.list_budget<-lapply(MonkeyIDs_budget, function(x){pre_focals_budget[pre_focals_budget[["FOCAL.ID"]] == x, ]})
head(monkey.list_budget)

#3. Create new list of tibbles to look at behavior occurrences per focal
monkey.list_pre_budget<-
  monkey.list_budget %>%
  purrr::map(~mutate(.,number.focals= n_distinct(focal.reference))) %>%
  purrr::map(~group_by(.,focal.reference,behavior.category)) %>% #grouping by focal# & behavior category
  purrr::map(~summarize(.,count=n())) %>%
  purrr::map(~group_by(.,behavior.category)) %>%
  purrr::map(~summarize(.,total= sum(count)))
head(monkey.list_pre_budget)
#This works fine and does exactly what I want but then I can't figure out how to divide the "total" column by the number of focals per individual
  
monkey.list_pre_budget<-
  monkey.list_pre_budget %>%
  purrr::map(~mutate(.,Mean = (total/monkey.list_pre_budget$number.focals)))
  
purrr::map(~mutate(.,count= n())) %>%
  purrr::map(~group_by(.,behavior.category)) %>%
  
  
  #purrr::map(~summarize(.,count=n())) %>% #counting up behavior categories
  purrr::map(~select(.,number.focals))
  purrr::map(~group_by(.,behavior.category)) %>%
  purrr::map(~summarize(.,total= sum(count)))
  #purrr::map(~group_by(.,behavior.category)) %>% #now grouping by the behavior categories only
  #purrr::map(~summarise(.,Mean= mean(count))) #getting the mean for each category per focal
names(monkey.list_pre_budget)<- MonkeyIDs_budget
head(monkey.list_pre_budget)

monkey.list_pre_budget<-
  monkey.list_pre_budget %>%
  purrr::map(~mutate(.,Mean = (total/(n_distinct(monkey.list_pre_budget$focal.reference)))))

#4. Combining all the tibbles into one
monkey_pre_budget<- monkey.list_pre_budget %>% 
  purrr::reduce(dplyr::full_join) %>% #using full_join since all columns are the same across tibbles
  mutate(release = "pre") %>% #creating a new column called "release"
  group_by(behavior.category) %>%
  summarise(Mean= mean(Mean))
monkey_pre_budget
```

```{r}
bp<- ggplot(monkey.list_pre_budget, aes(x="", y=Mean, fill=behavior.category))+
geom_bar(width = 1, stat = "identity")
bp

pie <- bp + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(Mean*100), "%")), position = position_stack(vjust = 0.5))
pie

blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )

library(scales)
pie + scale_fill_brewer("Blues") + blank_theme +
  theme(axis.text.x=element_blank())+
  geom_text(aes(y = Mean/8 + c(0, cumsum(Mean)[-length(Mean)]), 
                label = percent(Mean/100)), size=2)


```

