---
title: "2016 release"
output: html_document
---

install.packages("curl")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("gplots")
install.packages("corrplot")
install.packages("ggplot2")
install.packages("vegan")
install.packages("data.table")
install.packages("tidyr")
install.packages("formattable")

```{r}
library(curl)
library(dplyr)
library(tidyverse)
library(gplots)
library(corrplot)
library(ggplot2)
library(vegan)
library(data.table)
library(tidyr)
library(formattable)
```

## Importing the pre-focals datasheet
```{r}
pre_focals<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/2016_pre-release_FOCALS.csv")
pre_focals<- read.csv(pre_focals, header = T)
head(pre_focals)
nrow(pre_focals)
summary(pre_focals)
names(pre_focals)
unique(pre_focals$FOCAL.ID)
n_distinct(pre_focals$FOCAL.ID)
unique(pre_focals$AGE)
sort(table(pre_focals$BEHAVIOUR), decreasing = TRUE)
n_distinct(pre_focals$FOCAL.REFERENCE.CODE) #1012 total focals
```

## Creating an Age/Sex plot
```{r}
age_sex_plot <- ggplot(pre_focals, aes(SEX, group = AGE)) + 
          geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
          scale_y_continuous(labels=scales::percent) +
          ylab("relative frequencies") +
          facet_grid(~AGE)
age_sex_plot #There are more female than male adults, almost all subadults are males, and more juveniles are females... overall more females than males 
```

## POPs
## Filtering out Pops from the pre-focals dataset
```{r}
Pops<- pre_focals %>% filter(FOCAL.ID == "PO")
head(Pops)
n_distinct(Pops$FOCAL.REFERENCE.CODE) #56 distinct focals
unique(Pops$FOCAL.REFERENCE.CODE)
```

### Testing to see if I get the same SD value if I 1) do three focals separately and then find the average SD and 2)combine all three focals into one df and then find the SD for that 
```{r}
#These three chunks are each focal separately
Pops_behaviorcount<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10101") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount

Pops_behaviorcount2<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10407") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount2

Pops_behaviorcount3<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "20109") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount3

#Finding SD for each focal
SD1<- diversity(Pops_behaviorcount$count)
SD2<- diversity(Pops_behaviorcount2$count)
SD3<- diversity(Pops_behaviorcount3$count)

SD1
SD2
SD3
SDs<- c(SD1,SD2,SD3) #combining the focals
AvgSD<- mean(SDs)
AvgSD #1.145476

#Now I'm filtering out all 3 focals so that it counts the behaviors across all three
Pops_behaviorcount4<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE %in% c("10101","10407","20109")) %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount4

#finding the SD for this one df
SD4<- diversity(Pops_behaviorcount4$count)
SD4 #1.307044

#The SD values are DIFFERENT for the two methods! 
```

### Getting one SD value for all of Pops 56 focals... I don't think this is a reliable value?
```{r}
Pops_behaviorcountTEST2<- Pops %>% 
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcountTEST2

DIVTEST<- diversity(Pops_behaviorcountTEST2$count)
DIVTEST #1.459
```

### Playing around with the two behavior columns for one of Pop's focals, I need to somehow incorporate the behaviors from the Behavior.2 column into my SD calculations (don't know how to do this yet)
```{r}
Pops_behaviorcount2<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10407") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount2

Pops_behaviorcount3<- Pops %>% 
  filter(FOCAL.REFERENCE.CODE == "10407") %>%
  group_by(BEHAVIOUR.2) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount3
```

### Practicing running SD calculations by "hand" first
```{r}
sum(Pops_behaviorcount$count) #20
pI <- Pops_behaviorcount$count/sum(Pops_behaviorcount$count)
pI
H.I <- -sum(pI*log(pI))
H.I #0.7999
test<- diversity(Pops_behaviorcount$count)
test #0.7999, they match
```

### Since I can't find SD for each of the 56 focals (don't know how to do it at least), I'm going to find a SD for each month during pre-release observations
```{r}
#Month 6
Pops_behaviorcount_month6<- Pops %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount_month6

Pop_month6_SD<- diversity(Pops_behaviorcount_month6$count)
Pop_month6_SD

#Month 7
Pops_behaviorcount_month7<- Pops %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount_month7

Pop_month7_SD<- diversity(Pops_behaviorcount_month7$count)
Pop_month7_SD

#Month 8
Pops_behaviorcount_month8<- Pops %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount_month8

Pop_month8_SD<- diversity(Pops_behaviorcount_month8$count)
Pop_month8_SD

#Month 9
Pops_behaviorcount_month9<- Pops %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Pops_behaviorcount_month9

Pop_month9_SD<- diversity(Pops_behaviorcount_month9$count)
Pop_month9_SD
```

### Averaging the SDs across Pop's 4 months pre-release
```{r}
Pops_pre_allSDs<- c(Pop_month6_SD,Pop_month7_SD,Pop_month8_SD,Pop_month9_SD)
Pops_pre_allSDs
Pop_pre_avgSDs<- mean(Pops_pre_allSDs)
Pop_pre_avgSDs #1.377789
```

## ZIP
## Filtering out Zip from the pre-focals dataset
```{r}
Zip<- pre_focals %>% filter(FOCAL.ID == "ZI")
n_distinct(Zip$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Zip during pre-release observations
```{r}
#Month 6
Zip_behaviorcount_month6<- Zip %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Zip_behaviorcount_month6

Zip_month6_SD<- diversity(Zip_behaviorcount_month6$count)
Zip_month6_SD

#Month 7
Zip_behaviorcount_month7<- Zip %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Zip_behaviorcount_month7

Zip_month7_SD<- diversity(Zip_behaviorcount_month7$count)
Zip_month7_SD

#Month 8
Zip_behaviorcount_month8<- Zip %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Zip_behaviorcount_month8

Zip_month8_SD<- diversity(Zip_behaviorcount_month8$count)
Zip_month8_SD

#Month 9
Zip_behaviorcount_month9<- Zip %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Zip_behaviorcount_month9

Zip_month9_SD<- diversity(Zip_behaviorcount_month9$count)
Zip_month9_SD
```

### Averaging the SDs across Zip's 4 months pre-release
```{r}
Zip_pre_allSDs<- c(Zip_month6_SD,Zip_month7_SD,Zip_month8_SD,Zip_month9_SD)
Zip_pre_avgSDs<- mean(Zip_pre_allSDs)
Zip_pre_avgSDs
```

## JACK
## Filtering out Jack from the pre-focals dataset
```{r}
Jack<- pre_focals %>% filter(FOCAL.ID == "JA")
n_distinct(Jack$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Jack during pre-release observations
```{r}
#Month 6
Jack_behaviorcount_month6<- Jack %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Jack_behaviorcount_month6

Jack_month6_SD<- diversity(Jack_behaviorcount_month6$count)
Jack_month6_SD

#Month 7
Jack_behaviorcount_month7<- Jack %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Jack_behaviorcount_month7

Jack_month7_SD<- diversity(Jack_behaviorcount_month7$count)
Jack_month7_SD

#Month 8
Jack_behaviorcount_month8<- Jack %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Jack_behaviorcount_month8

Jack_month8_SD<- diversity(Jack_behaviorcount_month8$count)
Jack_month8_SD

#Month 9
Jack_behaviorcount_month9<- Jack %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Jack_behaviorcount_month9

Jack_month9_SD<- diversity(Jack_behaviorcount_month9$count)
Jack_month9_SD
```

### Averaging the SDs across Jack's 4 months pre-release
```{r}
Jack_pre_allSDs<- c(Jack_month6_SD,Jack_month7_SD,Jack_month8_SD,Jack_month9_SD)
Jack_pre_avgSDs<- mean(Jack_pre_allSDs)
Jack_pre_avgSDs
```

## BLUE
## Filtering out Blue from the pre-focals dataset
```{r}
Blue<- pre_focals %>% filter(FOCAL.ID == "BL")
n_distinct(Blue$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Blue during pre-release observations
```{r}
#Month 6
Blue_behaviorcount_month6<- Blue %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Blue_behaviorcount_month6

Blue_month6_SD<- diversity(Blue_behaviorcount_month6$count)
Blue_month6_SD

#Month 7
Blue_behaviorcount_month7<- Blue %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Blue_behaviorcount_month7

Blue_month7_SD<- diversity(Blue_behaviorcount_month7$count)
Blue_month7_SD

#Month 8
Blue_behaviorcount_month8<- Blue %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Blue_behaviorcount_month8

Blue_month8_SD<- diversity(Blue_behaviorcount_month8$count)
Blue_month8_SD

#Month 9
Blue_behaviorcount_month9<- Blue %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Blue_behaviorcount_month9

Blue_month9_SD<- diversity(Blue_behaviorcount_month9$count)
Blue_month9_SD
```

### Averaging the SDs across Blue's 4 months pre-release
```{r}
Blue_pre_allSDs<- c(Blue_month6_SD,Blue_month7_SD,Blue_month8_SD,Blue_month9_SD)
Blue_pre_avgSDs<- mean(Blue_pre_allSDs)
Blue_pre_avgSDs
```

## ALEX
## Filtering out Alex from the pre-focals dataset
```{r}
Alex<- pre_focals %>% filter(FOCAL.ID == "AL")
n_distinct(Alex$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Alex during pre-release observations
```{r}
#Month 6
Alex_behaviorcount_month6<- Alex %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Alex_behaviorcount_month6

Alex_month6_SD<- diversity(Alex_behaviorcount_month6$count)
Alex_month6_SD

#Month 7
Alex_behaviorcount_month7<- Alex %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Alex_behaviorcount_month7

Alex_month7_SD<- diversity(Alex_behaviorcount_month7$count)
Alex_month7_SD

#Month 8
Alex_behaviorcount_month8<- Alex %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Alex_behaviorcount_month8

Alex_month8_SD<- diversity(Alex_behaviorcount_month8$count)
Alex_month8_SD

#Month 9
Alex_behaviorcount_month9<- Alex %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Alex_behaviorcount_month9

Alex_month9_SD<- diversity(Alex_behaviorcount_month9$count)
Alex_month9_SD
```

### Averaging the SDs across Alex's 4 months pre-release
```{r}
Alex_pre_allSDs<- c(Alex_month6_SD,Alex_month7_SD,Alex_month8_SD,Alex_month9_SD)
Alex_pre_avgSDs<- mean(Alex_pre_allSDs)
Alex_pre_avgSDs
```

## BART
## Filtering out Bart from the pre-focals dataset
```{r}
Bart<- pre_focals %>% filter(FOCAL.ID == "BA")
n_distinct(Bart$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Bart during pre-release observations
```{r}
#Month 6
Bart_behaviorcount_month6<- Bart %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bart_behaviorcount_month6

Bart_month6_SD<- diversity(Bart_behaviorcount_month6$count)
Bart_month6_SD

#Month 7
Bart_behaviorcount_month7<- Bart %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bart_behaviorcount_month7

Bart_month7_SD<- diversity(Bart_behaviorcount_month7$count)
Bart_month7_SD

#Month 8
Bart_behaviorcount_month8<- Bart %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bart_behaviorcount_month8

Bart_month8_SD<- diversity(Bart_behaviorcount_month8$count)
Bart_month8_SD

#Month 9
Bart_behaviorcount_month9<- Bart %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bart_behaviorcount_month9

Bart_month9_SD<- diversity(Bart_behaviorcount_month9$count)
Bart_month9_SD
```

## NEVILLE
## Filtering out Nev from the pre-focals dataset
```{r}
Nev<- pre_focals %>% filter(FOCAL.ID == "NE")
n_distinct(Nev$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Neville during pre-release observations
```{r}
#Month 6
Nev_behaviorcount_month6<- Nev %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Nev_behaviorcount_month6

Nev_month6_SD<- diversity(Nev_behaviorcount_month6$count)
Nev_month6_SD

#Month 7
Nev_behaviorcount_month7<- Nev %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Nev_behaviorcount_month7

Nev_month7_SD<- diversity(Nev_behaviorcount_month7$count)
Nev_month7_SD

#Month 8
Nev_behaviorcount_month8<- Nev %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Nev_behaviorcount_month8

Nev_month8_SD<- diversity(Nev_behaviorcount_month8$count)
Nev_month8_SD

#Month 9
Nev_behaviorcount_month9<- Nev %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Nev_behaviorcount_month9

Nev_month9_SD<- diversity(Nev_behaviorcount_month9$count)
Nev_month9_SD
```

### Averaging the SDs across Nev's 4 months pre-release
```{r}
Nev_pre_allSDs<- c(Nev_month6_SD,Nev_month7_SD,Nev_month8_SD,Nev_month9_SD)
Nev_pre_avgSDs<- mean(Nev_pre_allSDs)
Nev_pre_avgSDs
```

## AUGUSTINE
## Filtering out Aug from the pre-focals dataset
```{r}
Aug<- pre_focals %>% filter(FOCAL.ID == "AU")
n_distinct(Aug$FOCAL.REFERENCE.CODE) #58 distinct focals
```

### Getting SD for each month for Aug during pre-release observations
```{r}
#Month 6
Aug_behaviorcount_month6<- Aug %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Aug_behaviorcount_month6

Aug_month6_SD<- diversity(Aug_behaviorcount_month6$count)
Aug_month6_SD

#Month 7
Aug_behaviorcount_month7<- Aug %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Aug_behaviorcount_month7

Aug_month7_SD<- diversity(Aug_behaviorcount_month7$count)
Aug_month7_SD

#Month 8
Aug_behaviorcount_month8<- Aug %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Aug_behaviorcount_month8

Aug_month8_SD<- diversity(Aug_behaviorcount_month8$count)
Aug_month8_SD

#Month 9
Aug_behaviorcount_month9<- Aug %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Aug_behaviorcount_month9

Aug_month9_SD<- diversity(Aug_behaviorcount_month9$count)
Aug_month9_SD
```

### Averaging the SDs across Aug's 4 months pre-release
```{r}
Aug_pre_allSDs<- c(Aug_month6_SD,Aug_month7_SD,Aug_month8_SD,Aug_month9_SD)
Aug_pre_avgSDs<- mean(Aug_pre_allSDs)
Aug_pre_avgSDs
```

## AMY
## Filtering out Amy from the pre-focals dataset
```{r}
Amy<- pre_focals %>% filter(FOCAL.ID == "AM")
n_distinct(Amy$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Amy during pre-release observations
```{r}
#Month 6
Amy_behaviorcount_month6<- Amy %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Amy_behaviorcount_month6

Amy_month6_SD<- diversity(Amy_behaviorcount_month6$count)
Amy_month6_SD

#Month 7
Amy_behaviorcount_month7<- Amy %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Amy_behaviorcount_month7

Amy_month7_SD<- diversity(Amy_behaviorcount_month7$count)
Amy_month7_SD

#Month 8
Amy_behaviorcount_month8<- Amy %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Amy_behaviorcount_month8

Amy_month8_SD<- diversity(Amy_behaviorcount_month8$count)
Amy_month8_SD

#Month 9
Amy_behaviorcount_month9<- Amy %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Amy_behaviorcount_month9

Amy_month9_SD<- diversity(Amy_behaviorcount_month9$count)
Amy_month9_SD
```

### Averaging the SDs across Amy's 4 months pre-release
```{r}
Amy_pre_allSDs<- c(Amy_month6_SD,Amy_month7_SD,Amy_month8_SD,Amy_month9_SD)
Amy_pre_avgSDs<- mean(Amy_pre_allSDs)
Amy_pre_avgSDs
```

## MAY
## Filtering out May from the pre-focals dataset
```{r}
May<- pre_focals %>% filter(FOCAL.ID == "MA")
n_distinct(May$FOCAL.REFERENCE.CODE) #57 distinct focals
```

### Getting SD for each month for May during pre-release observations
```{r}
#Month 6
May_behaviorcount_month6<- May %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
May_behaviorcount_month6

May_month6_SD<- diversity(May_behaviorcount_month6$count)
May_month6_SD

#Month 7
May_behaviorcount_month7<- May %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
May_behaviorcount_month7

May_month7_SD<- diversity(May_behaviorcount_month7$count)
May_month7_SD

#Month 8
May_behaviorcount_month8<- May %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
May_behaviorcount_month8

May_month8_SD<- diversity(May_behaviorcount_month8$count)
May_month8_SD

#Month 9
May_behaviorcount_month9<- May %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
May_behaviorcount_month9

May_month9_SD<- diversity(May_behaviorcount_month9$count)
May_month9_SD
```

### Averaging the SDs across May's 4 months pre-release
```{r}
May_pre_allSDs<- c(May_month6_SD,May_month7_SD,May_month8_SD,May_month9_SD)
May_pre_avgSDs<- mean(May_pre_allSDs)
May_pre_avgSDs
```

## TONI
## Filtering out Toni from the pre-focals dataset
```{r}
Toni<- pre_focals %>% filter(FOCAL.ID == "TO")
n_distinct(Toni$FOCAL.REFERENCE.CODE) #55 distinct focals
```

### Getting SD for each month for Toni during pre-release observations
```{r}
#Month 6
Toni_behaviorcount_month6<- Toni %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Toni_behaviorcount_month6

Toni_month6_SD<- diversity(Toni_behaviorcount_month6$count)
Toni_month6_SD

#Month 7
Toni_behaviorcount_month7<- Toni %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Toni_behaviorcount_month7

Toni_month7_SD<- diversity(Toni_behaviorcount_month7$count)
Toni_month7_SD

#Month 8
Toni_behaviorcount_month8<- Toni %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Toni_behaviorcount_month8

Toni_month8_SD<- diversity(Toni_behaviorcount_month8$count)
Toni_month8_SD

#Month 9
Toni_behaviorcount_month9<- Toni %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Toni_behaviorcount_month9

Toni_month9_SD<- diversity(Toni_behaviorcount_month9$count)
Toni_month9_SD
```

### Averaging the SDs across Toni's 4 months pre-release
```{r}
Toni_pre_allSDs<- c(Toni_month6_SD,Toni_month7_SD,Toni_month8_SD,Toni_month9_SD)
Toni_pre_avgSDs<- mean(Toni_pre_allSDs)
Toni_pre_avgSDs
```

## BOO
## Filtering out Boo from the pre-focals dataset
```{r}
Boo<- pre_focals %>% filter(FOCAL.ID == "BO")
n_distinct(Boo$FOCAL.REFERENCE.CODE) #55 distinct focals
```

### Getting SD for each month for Boo during pre-release observations
```{r}
#Month 6
Boo_behaviorcount_month6<- Boo %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Boo_behaviorcount_month6

Boo_month6_SD<- diversity(Boo_behaviorcount_month6$count)
Boo_month6_SD

#Month 7
Boo_behaviorcount_month7<- Boo %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Boo_behaviorcount_month7

Boo_month7_SD<- diversity(Boo_behaviorcount_month7$count)
Boo_month7_SD

#Month 8
Boo_behaviorcount_month8<- Boo %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Boo_behaviorcount_month8

Boo_month8_SD<- diversity(Boo_behaviorcount_month8$count)
Boo_month8_SD

#Month 9
Boo_behaviorcount_month9<- Boo %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Boo_behaviorcount_month9

Boo_month9_SD<- diversity(Boo_behaviorcount_month9$count)
Boo_month9_SD
```

### Averaging the SDs across Boo's 4 months pre-release
```{r}
Boo_pre_allSDs<- c(Boo_month6_SD,Boo_month7_SD,Boo_month8_SD,Boo_month9_SD)
Boo_pre_avgSDs<- mean(Boo_pre_allSDs)
Boo_pre_avgSDs
```

## BIG MAMA
## Filtering out Bgm from the pre-focals dataset
```{r}
Bgm<- pre_focals %>% filter(FOCAL.ID == "BM")
n_distinct(Bgm$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Bgm during pre-release observations
```{r}
#Month 6
Bgm_behaviorcount_month6<- Bgm %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bgm_behaviorcount_month6

Bgm_month6_SD<- diversity(Bgm_behaviorcount_month6$count)
Bgm_month6_SD

#Month 7
Bgm_behaviorcount_month7<- Bgm %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bgm_behaviorcount_month7

Bgm_month7_SD<- diversity(Bgm_behaviorcount_month7$count)
Bgm_month7_SD

#Month 8
Bgm_behaviorcount_month8<- Bgm %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bgm_behaviorcount_month8

Bgm_month8_SD<- diversity(Bgm_behaviorcount_month8$count)
Bgm_month8_SD

#Month 9
Bgm_behaviorcount_month9<- Bgm %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bgm_behaviorcount_month9

Bgm_month9_SD<- diversity(Bgm_behaviorcount_month9$count)
Bgm_month9_SD
```

### Averaging the SDs across Bgm's 4 months pre-release
```{r}
Bgm_pre_allSDs<- c(Bgm_month6_SD,Bgm_month7_SD,Bgm_month8_SD,Bgm_month9_SD)
Bgm_pre_avgSDs<- mean(Bgm_pre_allSDs)
Bgm_pre_avgSDs
```

## KOVU
## Filtering out Kovu from the pre-focals dataset
```{r}
Kovu<- pre_focals %>% filter(FOCAL.ID == "KO")
n_distinct(Kovu$FOCAL.REFERENCE.CODE) #55 distinct focals
```

### Getting SD for each month for Kovu during pre-release observations
```{r}
#Month 6
Kovu_behaviorcount_month6<- Kovu %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Kovu_behaviorcount_month6

Kovu_month6_SD<- diversity(Kovu_behaviorcount_month6$count)
Kovu_month6_SD

#Month 7
Kovu_behaviorcount_month7<- Kovu %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Kovu_behaviorcount_month7

Kovu_month7_SD<- diversity(Kovu_behaviorcount_month7$count)
Kovu_month7_SD

#Month 8
Kovu_behaviorcount_month8<- Kovu %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Kovu_behaviorcount_month8

Kovu_month8_SD<- diversity(Kovu_behaviorcount_month8$count)
Kovu_month8_SD

#Month 9
Kovu_behaviorcount_month9<- Kovu %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Kovu_behaviorcount_month9

Kovu_month9_SD<- diversity(Kovu_behaviorcount_month9$count)
Kovu_month9_SD
```

### Averaging the SDs across Kovu's 4 months pre-release
```{r}
Kovu_pre_allSDs<- c(Kovu_month6_SD,Kovu_month7_SD,Kovu_month8_SD,Kovu_month9_SD)
Kovu_pre_avgSDs<- mean(Kovu_pre_allSDs)
Kovu_pre_avgSDs
```

## EDDY
## Filtering out Eddy from the pre-focals dataset
```{r}
Ed<- pre_focals %>% filter(FOCAL.ID == "ED")
n_distinct(Ed$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Eddy during pre-release observations
```{r}
#Month 6
Ed_behaviorcount_month6<- Ed %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Ed_behaviorcount_month6

Ed_month6_SD<- diversity(Ed_behaviorcount_month6$count)
Ed_month6_SD

#Month 7
Ed_behaviorcount_month7<- Ed %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Ed_behaviorcount_month7

Ed_month7_SD<- diversity(Ed_behaviorcount_month7$count)
Ed_month7_SD

#Month 8
Ed_behaviorcount_month8<- Ed %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Ed_behaviorcount_month8

Ed_month8_SD<- diversity(Ed_behaviorcount_month8$count)
Ed_month8_SD

#Month 9
Ed_behaviorcount_month9<- Ed %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Ed_behaviorcount_month9

Ed_month9_SD<- diversity(Ed_behaviorcount_month9$count)
Ed_month9_SD
```

### Averaging the SDs across Eddy's 4 months pre-release
```{r}
Ed_pre_allSDs<- c(Ed_month6_SD,Ed_month7_SD,Ed_month8_SD,Ed_month9_SD)
Ed_pre_avgSDs<- mean(Ed_pre_allSDs)
Ed_pre_avgSDs
```

## TINKER
## Filtering out Tinker from the pre-focals dataset
```{r}
Tink<- pre_focals %>% filter(FOCAL.ID == "TI")
n_distinct(Tink$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Tinker during pre-release observations
```{r}
#Month 6
Tink_behaviorcount_month6<- Tink %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Tink_behaviorcount_month6

Tink_month6_SD<- diversity(Tink_behaviorcount_month6$count)
Tink_month6_SD

#Month 7
Tink_behaviorcount_month7<- Tink %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Tink_behaviorcount_month7

Tink_month7_SD<- diversity(Tink_behaviorcount_month7$count)
Tink_month7_SD

#Month 8
Tink_behaviorcount_month8<- Tink %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Tink_behaviorcount_month8

Tink_month8_SD<- diversity(Tink_behaviorcount_month8$count)
Tink_month8_SD

#Month 9
Tink_behaviorcount_month9<- Tink %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Tink_behaviorcount_month9

Tink_month9_SD<- diversity(Tink_behaviorcount_month9$count)
Tink_month9_SD
```

### Averaging the SDs across Tinker's 4 months pre-release
```{r}
Tink_pre_allSDs<- c(Tink_month6_SD,Tink_month7_SD,Tink_month8_SD,Tink_month9_SD)
Tink_pre_avgSDs<- mean(Tink_pre_allSDs)
Tink_pre_avgSDs
```

## BATMAN
## Filtering out Batman from the pre-focals dataset
```{r}
Bat<- pre_focals %>% filter(FOCAL.ID == "BT")
n_distinct(Bat$FOCAL.REFERENCE.CODE) #56 distinct focals
```

### Getting SD for each month for Batman during pre-release observations
```{r}
#Month 6
Bat_behaviorcount_month6<- Bat %>% 
  filter(MONTH == "6") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bat_behaviorcount_month6

Bat_month6_SD<- diversity(Bat_behaviorcount_month6$count)
Bat_month6_SD

#Month 7
Bat_behaviorcount_month7<- Bat %>% 
  filter(MONTH == "7") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bat_behaviorcount_month7

Bat_month7_SD<- diversity(Bat_behaviorcount_month7$count)
Bat_month7_SD

#Month 8
Bat_behaviorcount_month8<- Bat %>% 
  filter(MONTH == "8") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bat_behaviorcount_month8

Bat_month8_SD<- diversity(Bat_behaviorcount_month8$count)
Bat_month8_SD

#Month 9
Bat_behaviorcount_month9<- Bat %>% 
  filter(MONTH == "9") %>%
  group_by(BEHAVIOUR) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
Bat_behaviorcount_month9

Bat_month9_SD<- diversity(Bat_behaviorcount_month9$count)
Bat_month9_SD
```

### Averaging the SDs across Batman's 4 months pre-release
```{r}
Bat_pre_allSDs<- c(Bat_month6_SD,Bat_month7_SD,Bat_month8_SD,Bat_month9_SD)
Bat_pre_avgSDs<- mean(Bat_pre_allSDs)
Bat_pre_avgSDs
```

### Creating a table with all of average Shannon Diversity Indices for each individual
```{r}
pre_SDs_table<- matrix(c(Pop_pre_avgSDs, Zip_pre_avgSDs, Jack_pre_avgSDs, Blue_pre_avgSDs, Alex_pre_avgSDs, Bart_pre_avgSDs, Nev_pre_avgSDs, Aug_pre_avgSDs, Amy_pre_avgSDs, May_pre_avgSDs, Toni_pre_avgSDs, Boo_pre_avgSDs, Bgm_pre_avgSDs, Kovu_pre_avgSDs, Ed_pre_avgSDs, Tink_pre_avgSDs, Bat_pre_avgSDs), ncol = 1, byrow = 17)
pre_SDs_table
colnames(pre_SDs_table) <- c("Shannon Diversity Index")
rownames(pre_SDs_table) <- c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Bgm", "Kovu", "Eddy", "Tinker", "Batman")
pre_SDs_table<- as.table(pre_SDs_table)
#table(pre_SDs_table) %>% 
        #as.data.frame() %>% 
        #arrange(desc(Freq))
pre_SDs_table

x<- barplot(pre_SDs_table, beside = TRUE, space = c(0.25,1), xlab = "", ylab = "Shannon Diversity Index", names.arg = c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman"), ylim = c(0,2.5), cex.names = 0.8, las=2, col="#339999")#I don't love this barplot visual, I can't figure out how to turn the x-axis labels on an angle using a table
x
```
### Saving the image as a jpeg
```{r}
pdf(file = "/New User/Desktop/SD Plot.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches

jpeg("SDplot.jpg", width = 900, height = 500)
x<- barplot(pre_SDs_table, beside = TRUE, space = c(0.25,1), xlab = "", ylab = "Shannon Diversity Index", names.arg = c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman"), ylim = c(0,2), cex.lab = 1.25, font.lab= 2, cex.names = 1, las=2, col="#339999")#I don't love this barplot visual, I can't figure out how to turn the x-axis labels on an angle using a table
x

dev.off()
```

## Creating a DF for the pre-release Shannon Diversity Index values so that I can use the barplot() function more easily
```{r}
Individual<- c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman") #my first column
SD<- c(Pop_pre_avgSDs, Zip_pre_avgSDs, Jack_pre_avgSDs, Blue_pre_avgSDs, Alex_pre_avgSDs, Bart_pre_avgSDs, Nev_pre_avgSDs, Aug_pre_avgSDs, Amy_pre_avgSDs, May_pre_avgSDs, Toni_pre_avgSDs, Boo_pre_avgSDs, Bgm_pre_avgSDs, Kovu_pre_avgSDs, Ed_pre_avgSDs, Tink_pre_avgSDs, Bat_pre_avgSDs) #My second column
BD_df<- data.frame(Individual,SD) #Turn into data frame
BD_df

row.names(BD_df)<- c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman") #I need to customize my rownames because with the code I use down below to turn the names on an angle I need to use the paste(rownames) function... otherwise they would just be 1,2,3...
row.names(BD_df)

par(mar = c(11, 7, 2, 2) + 0.2) #add room for the rotated labels

end_point = 0.5 + nrow(BD_df) + nrow(BD_df) - 1 #this is the line which does the trick (together with barplot "space = 1" parameter)

barplot(BD_df$SD, col = "#69b3a2", 
        main = "",
        ylab = "Shannon Diversity Index", ylim = c(0,2.5),
        xlab = "",
        space = 0.75)
text(seq(1.5, end_point, by = 2), par("usr")[3]-0.25, 
     srt = 60, adj = 1, xpd = TRUE, #rotate 60 degrees (srt = 60)
     labels = paste(rownames(BD_df)), cex = 0.75)
```

### Creating a table by hand of the confirmed dead/presumed dead/presumed emigrated for males/females/infants
```{r}
info<- matrix(c(7,1,0,4,11,3,3,0,3,0,2,0,21,4,5,4), ncol=4, byrow = T)
rownames(info)<- c("Male", "Female", "Infant", "Totals")
colnames(info)<- c("Released", "Confirmed Dead", "Presumed Dead", "Presumed Emigrated")
info<- as.table(info)
info
```

### Using formattable() function to create a nicer table
```{r}
summary<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/Summary%20Table.csv")
summary<- read.csv(summary, header = T) #Had to create an excel table and import as df to use formattable()

formattable(summary)

customGreen0 = "#DeF7E9"

customGreen = "#71CA97"

customRed = "#ff7f7f"

formattable(summary, 
            align =c("l","c","c","c","c"), 
            list(`Vervets` = formatter(
              "span", style = ~ style(color = "grey",font.weight = "bold"))
))
```


```{r}
n_distinct(Pops$POSITION.IN.CANOPY) #8 (blanks makes 9)
unique(Pops$POSITION.IN.CANOPY)
Pops
```


