---
title: "Activity Budgets"
output: html_document
---

```{r}
library(lme4)
```

## Activity Budgets PRE RELEASE
### Behavior Categories
```{r}
pre_focals_budget<- pre_focals %>%
  #filter(FOCAL.ID != "BT") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>% #gathering the two behavior columns
  filter(!behavior %in% c("AL","BL","PO")) %>% #removing these from the behavior column becuase they're mistakes
  mutate(behavior.category = #creating new column for behavior categories
           case_when(behavior %in% c("G+","G-","PR-","PR+","C","CL","N","NU","SU","PL","MA","MO") ~ as.character("Social"),
                     behavior %in% c("F","FE","FO") ~ as.character("Feeding"),
                     behavior %in% c("L","R") ~ as.character("Activity"),
                     behavior %in% c("V","PA") ~ as.character("Predator"),
                     behavior %in% c("A+","A-","TH+","TH-","MP+","MP-") ~ as.character("Dominance"),
                     behavior %in% c("YA","SC","SG","SM","PC") ~ as.character("Stress"),
                     behavior %in% c("PH","AH")~ as.character("Human"),
                     behavior %in% c("O","OS")~ as.character("Other")))
unique(pre_focals_budget$behavior.category)
pre_focals_budget %>% filter(behavior %in% c("PH","AH"))
```

### Mean Activity Budget Per Focal
```{r}
#Pops
Pops_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "PO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n()) %>%
  mutate(ID = "PO")

Pops_budget2<- Pops_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Pops_budget$focal.reference))))
Pops_budget2$total<- NULL

#Blue
Blue_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BL") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n()) %>%
  mutate(ID = "BL")

Blue_budget2<- Blue_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Blue_budget$focal.reference))))
Blue_budget2$total<- NULL

#Amy
Amy_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "AM") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n()) %>%
  mutate(ID = "AM")

Amy_budget2<- Amy_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Amy_budget$focal.reference))))
Amy_budget2$total<- NULL

#Alex
Alex_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "AL") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Alex_budget<- Alex_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Alex_budget$focal.reference))))
Alex_budget$total<- NULL

#Boo
Boo_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Boo_budget<- Boo_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Boo_budget$focal.reference))))
Boo_budget$total<- NULL

#Big Mama
Bgm_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BM") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Bgm_budget<- Bgm_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Bgm_budget$focal.reference))))
Bgm_budget$total<- NULL

#Augustine
Aug_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "AU") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Aug_budget<- Aug_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Aug_budget$focal.reference))))
Aug_budget$total<- NULL

#Bart
Bart_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Bart_budget<- Bart_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Bart_budget$focal.reference))))
Bart_budget$total<- NULL

#Eddy
Ed_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "ED") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Ed_budget<- Ed_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Ed_budget$focal.reference))))
Ed_budget$total<- NULL

#Jack
Jack_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "JA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Jack_budget<- Jack_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Jack_budget$focal.reference))))
Jack_budget$total<- NULL

#Tinker
Tink_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "TI") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Tink_budget<- Tink_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Tink_budget$focal.reference))))
Tink_budget$total<- NULL

#Kovu
Kovu_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "KO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Kovu_budget<- Kovu_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Kovu_budget$focal.reference))))
Kovu_budget$total<- NULL

#Mango
Mango_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "KO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Mango_budget<- Mango_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Mango_budget$focal.reference))))
Mango_budget$total<- NULL

#Neville
Nev_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "NE") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Nev_budget<- Nev_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Nev_budget$focal.reference))))
Nev_budget$total<- NULL

#Toni
Toni_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "TO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Toni_budget<- Toni_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Toni_budget$focal.reference))))
Toni_budget$total<- NULL

#May
May_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "MA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

May_budget<- May_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(May_budget$focal.reference))))
May_budget$total<- NULL

#Zip
Zip_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "ZI") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Zip_budget<- Zip_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Zip_budget$focal.reference))))
Zip_budget$total<- NULL
```

*Troop Mean Activity Budgets*
```{r}
pre_fullbudget<- Pops_budget %>%
  full_join(Blue_budget,by= "behavior.category") %>%
  full_join(Amy_budget,by= "behavior.category") %>%
  full_join(Alex_budget,by= "behavior.category") %>%
  full_join(Boo_budget,by= "behavior.category") %>%
  full_join(Bgm_budget,by= "behavior.category") %>%
  full_join(Aug_budget,by= "behavior.category") %>%
  full_join(Bart_budget,by= "behavior.category") %>%
  full_join(Ed_budget,by= "behavior.category") %>%
  full_join(Jack_budget,by= "behavior.category") %>%
  full_join(Tink_budget,by= "behavior.category") %>%
  full_join(Kovu_budget,by= "behavior.category") %>%
  full_join(Mango_budget,by= "behavior.category") %>%
  full_join(May_budget,by= "behavior.category") %>%
  full_join(Toni_budget,by= "behavior.category") %>%
  full_join(Nev_budget,by= "behavior.category") %>%
  full_join(Zip_budget,by= "behavior.category") %>%
  rowwise() %>% #this allows us to work with row-wise dataframe where we want to look at each row individually
  mutate(TroopMean = ((sum(c_across(2:18),na.rm=T))/17), #adding up all the means per category and dividing by 17 monkeys to get average
        release = "pre") #adding column to identify pre-release

pre_fullbudget<- pre_fullbudget[,c("behavior.category","TroopMean","release")] #these are the only columns I want to look at now
pre_fullbudget
```

### Activity Budget Bar Chart
```{r}
budget_plot<-ggplot(data=pre_fullbudget, aes(x=behavior.category, y=TroopMean)) +
  geom_bar(stat="identity")
budget_plot
```

### Batman vs Rehabilitant Males: Activity Budget 
```{r}
#Batman
Bat_budget<- pre_focals_budget %>%
  filter(FOCAL.ID == "BT") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())
Bat_budget

Bat_budget<- Bat_budget %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Bat_budget$focal.reference))))
Bat_budget$total<- NULL

Pops_budget_pred<- Pops_budget %>%
  filter(behavior.category == "Predator") %>%
  mutate(ID = "Pops")
Pops_budget_pred

Bat_budget_pred<- Bat_budget %>%
  filter(behavior.category == "Predator") %>%
  mutate(ID = "Batman")

Males_budget<- bind_rows(Pops_budget_pred, Bat_budget_pred)
Males_budget %>% filter(ID == "Pops")
```

### Trying to do Wilcoxon Test with Batman and Pops
```{r}
ggdensity(Males_budget$count)
ggqqplot(Males_budget$count)

ggboxplot(Males_budget, x = "ID", y = "count", 
          color = "ID", palette = c("#00AFBB", "#E7B800"),
          ylab = "count", xlab = "ID")

WT_PopsBat_pred<- wilcox.test(count ~ ID, data = Males_budget, paired = FALSE)
WT_PopsBat_pred
```

## Activity Budgets POST RELEASE
### Behavior Categories
```{r}
post_focals_budget<- post_focals %>%
  filter(!FOCAL.ID %in% c("CI","HO","PA")) %>%
  gather("Obs", "behavior", 20:21, na.rm = T) %>% #gathering the two behavior columns
  filter(!behavior %in% c("ED","TH","VL")) %>% #removing these from the behavior column becuase they're mistakes
  mutate(behavior.category = #creating new column for behavior categories
           case_when(behavior %in% c("G+","G-","PR-","PR+","C","CL","N","SU","PL","MA","MO") ~ as.character("Social"),
                     behavior %in% c("F","FE","FO") ~ as.character("Feeding"),
                     behavior %in% c("L","R") ~ as.character("Activity"),
                     behavior %in% c("V","PA") ~ as.character("Predator"),
                     behavior %in% c("A+","A-","TH+","TH-","MP+","MP-") ~ as.character("Dominance"),
                     behavior %in% c("YA","SC","SG","SM","PC") ~ as.character("Stress"),
                     behavior %in% c("PH","AH")~ as.character("Human"),
                     behavior %in% c("O","OS","oS")~ as.character("Other"))) %>%
  rename(focal.reference = FOCAL.REFERENCE.CODE)
unique(post_focals_budget$behavior.category)
```

### Mean Activity Budget Per Focal
```{r}
#Pops
Pops_budget3<- post_focals_budget %>%
  filter(FOCAL.ID == "PO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n()) %>%
  mutate(ID = "PO")

Pops_budget4<- Pops_budget3 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Pops_budget3$focal.reference))))
Pops_budget4$total<- NULL

#Blue
Blue_budget3<- post_focals_budget %>%
  filter(FOCAL.ID == "BL") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n()) %>%
  mutate(ID = "BL")

Blue_budget4<- Blue_budget3 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Blue_budget3$focal.reference))))
Blue_budget4$total<- NULL

#Amy
Amy_budget3<- post_focals_budget %>%
  filter(FOCAL.ID == "AM") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n()) %>%
  mutate(ID = "AM")

Amy_budget4<- Amy_budget3 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Amy_budget3$focal.reference))))
Amy_budget4$total<- NULL

#Alex
Alex_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "AL") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Alex_budget2<- Alex_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Alex_budget2$focal.reference))))
Alex_budget2$total<- NULL

#Boo
Boo_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "BO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Boo_budget2<- Boo_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Boo_budget2$focal.reference))))
Boo_budget2$total<- NULL

#Big Mama
Bgm_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "BM") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Bgm_budget2<- Bgm_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Bgm_budget2$focal.reference))))
Bgm_budget2$total<- NULL

#Augustine
Aug_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "AU") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Aug_budget2<- Aug_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Aug_budget2$focal.reference))))
Aug_budget2$total<- NULL

#Eddy
Ed_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "ED") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Ed_budget2<- Ed_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Ed_budget2$focal.reference))))
Ed_budget2$total<- NULL

#Jack
Jack_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "JA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Jack_budget2<- Jack_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Jack_budget2$focal.reference))))
Jack_budget2$total<- NULL

#Tinker
Tink_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "TI") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Tink_budget2<- Tink_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Tink_budget2$focal.reference))))
Tink_budget2$total<- NULL

#Kovu
Kovu_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "KO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Kovu_budget2<- Kovu_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Kovu_budget2$focal.reference))))
Kovu_budget2$total<- NULL

#Mango
Mango_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "KO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Mango_budget2<- Mango_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Mango_budget2$focal.reference))))
Mango_budget2$total<- NULL

#Neville
Nev_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "NE") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Nev_budget2<- Nev_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Nev_budget2$focal.reference))))
Nev_budget2$total<- NULL

#Toni
Toni_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "TO") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Toni_budget2<- Toni_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Toni_budget2$focal.reference))))
Toni_budget2$total<- NULL

#May
May_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "MA") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

May_budget2<- May_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(May_budget2$focal.reference))))
May_budget2$total<- NULL

#Zip
Zip_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "ZI") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Zip_budget2<- Zip_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Zip_budget2$focal.reference))))
Zip_budget2$total<- NULL

#Red
Red_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "RE") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Red_budget2<- Red_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Red_budget2$focal.reference))))
Red_budget2$total<- NULL

#Skittles
Sk_budget2<- post_focals_budget %>%
  filter(FOCAL.ID == "SK") %>%
  group_by(focal.reference,behavior.category) %>%
  summarize(count=n())

Sk_budget2<- Sk_budget2 %>%
  group_by(behavior.category) %>%
  summarize(total= sum(count)) %>%
  mutate(Mean = (total/(n_distinct(Sk_budget2$focal.reference))))
Sk_budget2$total<- NULL
```

*Troop Mean Activity Budgets*
```{r}
post_fullbudget<- Pops_budget2 %>%
  full_join(Blue_budget2,by= "behavior.category") %>%
  full_join(Amy_budget2,by= "behavior.category") %>%
  full_join(Alex_budget2,by= "behavior.category") %>%
  full_join(Boo_budget2,by= "behavior.category") %>%
  full_join(Bgm_budget2,by= "behavior.category") %>%
  full_join(Aug_budget2,by= "behavior.category") %>%
  full_join(Ed_budget2,by= "behavior.category") %>%
  full_join(Jack_budget2,by= "behavior.category") %>%
  full_join(Tink_budget2,by= "behavior.category") %>%
  full_join(Kovu_budget2,by= "behavior.category") %>%
  full_join(Mango_budget2,by= "behavior.category") %>%
  full_join(May_budget2,by= "behavior.category") %>%
  full_join(Toni_budget2,by= "behavior.category") %>%
  full_join(Nev_budget2,by= "behavior.category") %>%
  full_join(Zip_budget2,by= "behavior.category") %>%
  full_join(Red_budget2,by= "behavior.category") %>%
  full_join(Sk_budget2,by= "behavior.category") %>%
  rowwise() %>% #this allows us to work with row-wise dataframe where we want to look at each row individually
  mutate(TroopMean = ((sum(c_across(2:19),na.rm=T))/19), #adding up all the means per category and dividing by 19 monkeys to get average
        release= "post") #adding column to identify pre-release

post_fullbudget<- post_fullbudget[,c("behavior.category","TroopMean","release")] #these are the only columns I want to look at now
post_fullbudget
```

### MEAN Activity Budget Tables
```{r}
combined_budget_table<- bind_rows(pre_fullbudget, post_fullbudget)
combined_budget_table<- combined_budget_table %>%
  mutate(proportion= TroopMean/20)
combined_budget_table

full_budgetplot<- ggplot(data=combined_budget_table, aes(x=behavior.category, y=proportion, fill=release)) +
  geom_bar(position="dodge", stat="identity") +
  scale_fill_discrete(name = "Release Stage", labels = c("Post", "Pre")) +
  scale_y_continuous(labels=scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust =1)) +
  xlab("Activity Budget Category") +
  ylab("Mean Behavior Occurrence Per Focal")
full_budgetplot
```
RESULTS: This figure shows the mean percentage of behavior occurences in each cateogry per focal observation across release stage. 

### COUNT Activity Budget Tables (to use for LMM)
```{r}
#creating a full activity budget table for all individuals post-release 
post_ABtable_full<- bind_rows(Pops_budget3,Blue_budget3,Amy_budget3) %>% #this data is counts per focal in each focals, NOT mean count across all of the focals
  mutate(release = "2")
post_ABtable_full

#creating a full activity budget table for all individuals pre-release
pre_ABtable_full<- bind_rows(Pops_budget,Blue_budget,Amy_budget) %>% #this data is counts per focal in each focals, NOT mean count across all of the focals
  mutate(release = "1")
pre_ABtable_full

#combining the two together with sex and age
ABtable_full<- bind_rows(pre_ABtable_full,post_ABtable_full) %>%
  mutate(sex = 
           case_when(ID == "PO" ~ as.character("0"), #making these character so that I can make them factors with labels in the next step 
                     ID %in% c("BL","AM")~ as.character("1")),
         age = 
           case_when(ID %in% c("PO","BL","AM")~ as.character("1"))) %>%
  rename(counts = count)

unique(ABtable_full$ID)


#making sure all of my columns are factors 
ABtable_full <- within(ABtable_full, {
  sex <- factor(sex, levels = 0:1, labels = c("M", "F"))
  age <- factor(age, levels = 1, labels = "A")
  release <- factor(release, levels = 1:2, labels = c("pre","post"))
  ID <- factor(ID, levels= c("PO","BL","AM"))
})
str(ABtable_full)
```

*Predator ONLY table*
```{r}
#filtering out the predator category
ABtable_full_pred<- ABtable_full %>%
  filter(behavior.category == "Predator")
ABtable_full_pred

#Checking histogram
hist(log(ABtable_full_pred$count))

pred_lm = lm(count ~ release, data = ABtable_full_pred)
summary(pred_lm)

pred_mixed<- lmer(count ~ release + (1 | ID), data = ABtable_full_pred)
summary(pred_mixed)

confint(pred_mixed)
```

## Example datasets from R
```{r}
hdp <- read.csv("https://stats.idre.ucla.edu/stat/data/hdp.csv")
hdp

Dyestuff
str(Penicillin)
```

## Kruskal Wallis Test
### Pre-release
```{r}
ABtable_pre<- ABtable_full %>%
  filter(release == "pre", behavior.category != "Human")

group_by(ABtable_pre, behavior.category) %>%
  summarise(
    count = n(),
    mean = mean(counts, na.rm = TRUE),
    sd = sd(counts, na.rm = TRUE),
    median = median(counts, na.rm = TRUE),
    IQR = IQR(counts, na.rm = TRUE)
  )

ggboxplot(ABtable_pre, x = "behavior.category", y = "counts", 
          ylab = "Activity Budget Category", xlab = "Mean Count Per Focal")

kruskal.test(counts ~ behavior.category, data = ABtable_pre)
pairwise.wilcox.test(ABtable_pre$counts, ABtable_pre$behavior.category,
                 p.adjust.method = "BH")
```
RESULTS: There is a significant difference across behavior categories in the troops activity budget pre-release. 

### Post-Release
```{r}
ABtable_post<- ABtable_full %>%
  filter(release == "post") #behavior.category != "Human")
unique(ABtable_post$behavior.category)

group_by(ABtable_post, behavior.category) %>%
  summarise(
    count = n(),
    mean = mean(counts, na.rm = TRUE),
    sd = sd(counts, na.rm = TRUE),
    median = median(counts, na.rm = TRUE),
    IQR = IQR(counts, na.rm = TRUE)
  )

ggboxplot(ABtable_pre, x = "behavior.category", y = "counts", 
          ylab = "Activity Budget Category", xlab = "Mean Count Per Focal")

kruskal.test(counts ~ behavior.category, data = ABtable_pre)
pairwise.wilcox.test(ABtable_pre$counts, ABtable_pre$behavior.category,
                 p.adjust.method = "BH")
```



```{r}
group_by(ABtable_full, behavior.category) %>%
  summarise(
    count = n(),
    mean = mean(counts, na.rm = TRUE),
    sd = sd(counts, na.rm = TRUE),
    median = median(counts, na.rm = TRUE),
    IQR = IQR(counts, na.rm = TRUE)
  )

ggboxplot(ABtable_full, x = "behavior.category", y = "counts", 
          ylab = "Activity Budget Category", xlab = "Mean Count Per Focal")

ggline(ABtable_full, x = "behavior.category", y = "counts", 
       add = c("mean_se", "jitter"), 
       ylab = "Activity Budget Category", xlab = "Mean Count Per Focal")

kruskal.test(counts ~ behavior.category, data = ABtable_full)
pairwise.wilcox.test(ABtable_full$counts, ABtable_full$behavior.category,
                 p.adjust.method = "BH")
```
RESULTS: There is a significant difference across behavior categories in the troops activity budget both pre- and post-release. 
