---
title: "Behavior Matrix"
output: html_document
---


```{r}
install.packages("dplyr")
library(curl)
library(dplyr)
library(tidyverse)
```

### Importing the FULL pre-continuous dataset
```{r}
pre_cont<- curl("https://raw.githubusercontent.com/langley1/LWTdata2016/main/2016_pre-release_cont_FULL.csv")
pre_cont<- read.csv(pre_cont, header=T, na.strings=c(""," ","NA"))
```

## Code for Behavior Matrices

### Grooming Matrix+ (focal recieves grooming)
```{r}
library(tidyverse)
# 1. Create a character vector of all the monkey IDs in your dataset:
MonkeyIDs<-as.character(unique(pre_cont$FOCAL.ID))

# 2. Get a list of dataframes, subsetted by monkey ID:
monkey.prelim<-lapply(MonkeyIDs, function(x){pre_cont[pre_cont[["FOCAL.ID"]] == x, ]})
head(monkey.prelim)

# 3. Filter each by the behavior your want, group each by associate/recipient, and count behavior:
monkey_G.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR=="G+")) %>%
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_G.prelim) <- MonkeyIDs

# 4. Set up your pairwise combinations of interacting monkeys:
monkeylist<-list(actor=MonkeyIDs,recipient=MonkeyIDs) #create list of all possible actors/recipients
filt <- function(x, y) {x == y} #create function to filter out same-monkey pairs ("FZ grooms FZ")
combo <- monkeylist %>% cross_df(.,.filter=filt) #get the filtered combined list as a dataframe

combo_G<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_G.prelim2),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_G.prelim2[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

G1<-combo_G %>% 
  mutate(grooms = map2_int(
    actor, 
    recipient, 
    ~monkey_G.prelim2 %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
G1

grooming_matrix<-spread(G1,recipient,grooms) %>% column_to_rownames(var="actor") %>% data.matrix()
grooming_matrix
```

### Grooming Matrix- (focal grooms)
```{r}
monkey_G2.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR=="G-")) %>%
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_G2.prelim) <- MonkeyIDs

# 4. Set up your pairwise combinations of interacting monkeys:
monkeylist<-list(actor=MonkeyIDs,recipient=MonkeyIDs) #create list of all possible actors/recipients
filt <- function(x, y) {x == y} #create function to filter out same-monkey pairs ("FZ grooms FZ")
combo <- monkeylist %>% cross_df(.,.filter=filt) #get the filtered combined list as a dataframe

combo_G2<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_G2.prelim),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_G2.prelim[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

G2<-combo_G2 %>% 
  mutate(grooms = map2_int(
    actor, 
    recipient, 
    ~monkey_G2.prelim %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
G2

grooming_matrix2<-spread(G2,recipient,grooms) %>% column_to_rownames(var="actor") %>% data.matrix()
grooming_matrix2
```


### Making Place+ Matrix (focal displaces the other individual)
```{r}
monkey_MP.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR=="MP+")) %>%
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_MP.prelim) <- MonkeyIDs

monkeylist<-list(actor=MonkeyIDs,recipient=MonkeyIDs) #create list of all possible actors/recipients
filt <- function(x, y) {x == y} #create function to filter out same-monkey pairs ("FZ grooms FZ")
combo <- monkeylist %>% cross_df(.,.filter=filt) #get the filtered combined list as a dataframe

combo_MP<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_MP.prelim),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_MP.prelim[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

MP1<-combo_MP %>% 
  mutate(makeplace = map2_int(
    actor, 
    recipient, 
    ~monkey_MP.prelim %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
MP1

MP_matrix<-spread(MP1,recipient,makeplace) %>% column_to_rownames(var="actor") %>% data.matrix()
MP_matrix
```

### Making Place- Matrix (focal gets displaced)
```{r}
monkey_MP2.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR=="MP-")) %>%
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_MP2.prelim) <- MonkeyIDs

monkeylist<-list(actor=MonkeyIDs,recipient=MonkeyIDs) #create list of all possible actors/recipients
filt <- function(x, y) {x == y} #create function to filter out same-monkey pairs ("FZ grooms FZ")
combo <- monkeylist %>% cross_df(.,.filter=filt) #get the filtered combined list as a dataframe

combo_MP2<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_MP2.prelim),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_MP2.prelim[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

MP2<-combo_MP2 %>% 
  mutate(makeplace = map2_int(
    actor, 
    recipient, 
    ~monkey_MP2.prelim %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
MP2

MP_matrix2<-spread(MP2,recipient,makeplace) %>% column_to_rownames(var="actor") %>% data.matrix()
MP_matrix2
```

### Aggression+ Matrix with conflict outcome= 1 (focal acts aggressively or recieves aggression and WINS)
```{r}
monkey_A.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR %in% c("A+","A-"),conflict.outcome=="1")) %>% #adding in a second filter so that I can pull out interactions where the focal won
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_A.prelim) <- MonkeyIDs

combo_A<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_A.prelim),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_A.prelim[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

A1<-combo_A %>% 
  mutate(aggress = map2_int(
    actor, 
    recipient, 
    ~monkey_A.prelim %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
A1

A_matrix<-spread(A1,recipient,aggress) %>% column_to_rownames(var="actor") %>% data.matrix()
A_matrix
```

### Aggression Matrix with conflict outcome= 2 (focal acts aggressively or recieves aggression and LOSES)
```{r}
monkey_loseA.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR %in% c("A+","A-"),conflict.outcome=="2")) %>% #adding in a second filter so that I can pull out interactions where the focal loses
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_loseA.prelim) <- MonkeyIDs

combo_loseA<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_loseA.prelim),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_loseA.prelim[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

loseA1<-combo_loseA %>% 
  mutate(aggress = map2_int(
    actor, 
    recipient, 
    ~monkey_loseA.prelim %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
loseA1

loseA_matrix<-spread(loseA1,recipient,aggress) %>% column_to_rownames(var="actor") %>% data.matrix()
loseA_matrix
```

### Threat Matrix with conflict outcome= 1 (focal threatens or recieves threat and WINS)
```{r}
monkey_TH.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR %in% c("TH+","TH-"),conflict.outcome=="1")) %>% 
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_TH.prelim) <- MonkeyIDs

combo_TH<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_TH.prelim),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_TH.prelim[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

TH1<-combo_TH %>% 
  mutate(threaten = map2_int(
    actor, 
    recipient, 
    ~monkey_TH.prelim %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
TH1

TH_matrix<-spread(TH1,recipient,threaten) %>% column_to_rownames(var="actor") %>% data.matrix()
TH_matrix
```

### Threat Matrix with conflict outcome= 2 (focal threatens or recieves threat and LOSES)
```{r}
monkey_loseTH.prelim<-
  monkey.prelim %>%
  purrr::map(~filter(.,BEHAVIOUR %in% c("TH+","TH-"),conflict.outcome=="2")) %>% 
  purrr::map(~group_by(.,ASSOCIATION)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey_loseTH.prelim) <- MonkeyIDs

combo_loseTH<-
  combo %>%
  mutate(absent1 = map2_chr(
    actor,
    recipient,
    ~if_else(.x %in% names(monkey_loseTH.prelim),true="TRUE",false="FALSE"))) %>%
    mutate(absent2 = map2_chr(
    actor,
    recipient,
    ~if_else(.y %in% monkey_loseTH.prelim[[.x]]$ASSOCIATION,true="TRUE",false="FALSE"))) %>%
  filter(absent1 == "TRUE") %>%
  filter(absent2 == "TRUE") %>%
  select(-absent1,-absent2)

TH2<-combo_loseTH %>% 
  mutate(threaten = map2_int(
    actor, 
    recipient, 
    ~monkey_loseTH.prelim %>% pluck(.x) %>% filter(ASSOCIATION==.y) %>% as.data.frame(.) %>% .[,2]))
TH2

loseTH_matrix<-spread(TH2,recipient,threaten) %>% column_to_rownames(var="actor") %>% data.matrix()
loseTH_matrix
```

```{r}
MP_matrix #focal wins, displaces the other
MP_matrix2 #focal gets displaced
A_matrix #focal wins any aggression interaction
loseA_matrix #focal loses any aggression interaction
TH_matrix #focal wins any threat interaction
loseTH_matrix #focal loses any threat interaction

MP_matrixDF<- as.data.frame(MP_matrix)
MP_matrixDF <- cbind(IDs = rownames(MP_matrixDF), MP_matrixDF)
rownames(MP_matrixDF) <- NULL
MP_matrixDF

x<- tibble(MP_matrixDF) %>% rowwise() %>% mutate(total = sum(across(AL:ZI)))
x
MP_matrixDF
MP_matrixDF<- MP_matrixDF

```

