---
title: "Shannon Diversity"
output: html_document
---

### Purrr method for SD
```{r}
#PRE-RELEASE
# 1. Create a character vector of all the monkey IDs in your dataset:
MonkeyIDs_SD<-as.character(unique(pre_focals$FOCAL.ID))

# 2. Get a list of dataframes, subsetted by monkey ID:
monkey.prelim_SD<-lapply(MonkeyIDs_SD, function(x){pre_focals[pre_focals[["FOCAL.ID"]] == x, ]})
head(monkey.prelim_SD)

# 3. Filter each by the behavior your want, group each by associate/recipient, and count behavior:
monkey.prelim_SD2 <-
  monkey.prelim_SD %>% 
  purrr::map(~gather(.,"Obs", "behavior", 22:23, na.rm = T)) %>%
  purrr::map(~filter(.,!behavior %in% c("PO","AL","BL"))) %>%
  purrr::map(~group_by(.,MONTH,behavior)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey.prelim_SD2) <- MonkeyIDs_SD 
head(monkey.prelim_SD2)

#POST-RELEASE
# 1. Create a character vector of all the monkey IDs in your dataset:
MonkeyIDs_SD2<-as.character(unique(post_focals$FOCAL.ID))

#2. 
monkey.prelim_SD_post<-lapply(MonkeyIDs_SD, function(x){pre_focals[pre_focals[["FOCAL.ID"]] == x, ]})
head(monkey.prelim_SD)

# 3. Filter each by the behavior your want, group each by associate/recipient, and count behavior:
monkey.prelim_SD2 <-
  monkey.prelim_SD %>% 
  purrr::map(~gather(.,"Obs", "behavior", 22:23, na.rm = T)) %>%
  purrr::map(~filter(.,behavior != ("ED"))) %>%
  purrr::map(~group_by(.,MONTH,behavior)) %>%
  purrr::map(~summarize(.,count=n()))
names(monkey.prelim_SD2) <- MonkeyIDs_SD 
head(monkey.prelim_SD2)
```

# PRE-RELEASE
## Pops
### Finding Pops SD for EACH MONTH during pre-release observations
```{r}
#Month 6
Pops_behaviorcount_month6<- Pops %>% 
  filter(MONTH == "6") %>% #filtering out month "6"
  gather("Obs", "behavior", 22:23, na.rm = T) %>% #using the gather() function to create two new columns "obs" and "behavior" using columns 19 & 20 ("Behavior" & "Behavior.2"). This adjusts the rows and basically combines all of the behaviors into one column under "behavior"
  group_by(behavior) %>% #Now grouping by this new "behavior" column created by gather()
  summarize(count=n()) %>% #counting the different behaviors and arranging them in descending order
  arrange(desc(count))
Pops_behaviorcount_month6

Pop_month6_SD<- diversity(Pops_behaviorcount_month6$count) #finding Shannon Diversity for for month 6
Pop_month6_SD

#Month 7
Pops_behaviorcount_month7<- Pops %>% 
  filter(MONTH == "7") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month7

Pop_month7_SD<- diversity(Pops_behaviorcount_month7$count)
Pop_month7_SD

#Month 8
Pops_behaviorcount_month8<- Pops %>% 
  filter(MONTH == "8") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month8

Pop_month8_SD<- diversity(Pops_behaviorcount_month8$count)
Pop_month8_SD

#Month 9
Pops_behaviorcount_month9<- Pops %>% 
  filter(MONTH == "9") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month9

Pop_month9_SD<- diversity(Pops_behaviorcount_month9$count)
Pop_month9_SD

#Month 10
Pops_behaviorcount_month10<- Pops %>% 
  filter(MONTH == "10") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month10

Pop_month10_SD<- diversity(Pops_behaviorcount_month10$count)
Pop_month10_SD

#Month 11
Pops_behaviorcount_month11<- Pops %>% 
  filter(MONTH == "11") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month11

Pop_month11_SD<- diversity(Pops_behaviorcount_month11$count)
Pop_month11_SD

#Month 12
Pops_behaviorcount_month12<- Pops %>% 
  filter(MONTH == "12") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month12

Pop_month12_SD<- diversity(Pops_behaviorcount_month12$count)
Pop_month12_SD

#Month 1
Pops_behaviorcount_month1<- Pops %>% 
  filter(MONTH == "1") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month1

Pop_month1_SD<- diversity(Pops_behaviorcount_month1$count)
Pop_month1_SD

#Month 2
Pops_behaviorcount_month2<- Pops %>% 
  filter(MONTH == "2") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Pops_behaviorcount_month2

Pop_month2_SD<- diversity(Pops_behaviorcount_month2$count)
Pop_month2_SD
```

### Averaging the SDs across Pop's months pre-release
```{r}
Pops_pre_allSDs<- c(Pop_month6_SD,Pop_month7_SD,Pop_month8_SD,Pop_month9_SD, Pop_month10_SD,Pop_month11_SD,Pop_month12_SD,Pop_month1_SD,Pop_month2_SD)
Pop_pre_avgSDs<- mean(Pops_pre_allSDs)
#Pop_pre_avgSDs #1.153
```

## ZIP
## Filtering out Zip from the pre-focals dataset
```{r}
Zip<- pre_focals %>% filter(FOCAL.ID == "ZI")
```

### Getting SD for each month for Zip during pre-release observations
```{r}
#Month 6
Zip_behaviorcount_month6<- Zip %>% 
  filter(MONTH == "6") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month6

Zip_month6_SD<- diversity(Zip_behaviorcount_month6$count)
Zip_month6_SD

#Month 7
Zip_behaviorcount_month7<- Zip %>% 
  filter(MONTH == "7") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month7

Zip_month7_SD<- diversity(Zip_behaviorcount_month7$count)
Zip_month7_SD

#Month 8
Zip_behaviorcount_month8<- Zip %>% 
  filter(MONTH == "8") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month8

Zip_month8_SD<- diversity(Zip_behaviorcount_month8$count)
Zip_month8_SD

#Month 9
Zip_behaviorcount_month9<- Zip %>% 
  filter(MONTH == "9") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month9

Zip_month9_SD<- diversity(Zip_behaviorcount_month9$count)
Zip_month9_SD

#Month 10
Zip_behaviorcount_month10<- Zip %>% 
  filter(MONTH == "10") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month10

Zip_month10_SD<- diversity(Zip_behaviorcount_month10$count)
Zip_month10_SD

#Month 11
Zip_behaviorcount_month11<- Zip %>% 
  filter(MONTH == "11") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month11

Zip_month11_SD<- diversity(Zip_behaviorcount_month11$count)
Zip_month11_SD

#Month 12
Zip_behaviorcount_month12<- Zip %>% 
  filter(MONTH == "12") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month12

Zip_month12_SD<- diversity(Zip_behaviorcount_month12$count)
Zip_month12_SD

#Month 1
Zip_behaviorcount_month1<- Zip %>% 
  filter(MONTH == "1") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month1

Zip_month1_SD<- diversity(Zip_behaviorcount_month1$count)
Zip_month1_SD

#Month 2
Zip_behaviorcount_month2<- Zip %>% 
  filter(MONTH == "2") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Zip_behaviorcount_month2

Zip_month2_SD<- diversity(Zip_behaviorcount_month2$count)
Zip_month2_SD
```

### Averaging the SDs across Zip's 4 months pre-release
```{r}
Zip_pre_allSDs<- c(Zip_month6_SD,Zip_month7_SD,Zip_month8_SD,Zip_month9_SD,Zip_month10_SD,Zip_month11_SD,Zip_month12_SD,Zip_month1_SD,Zip_month2_SD)
Zip_pre_avgSDs<- mean(Zip_pre_allSDs)
#Zip_pre_avgSDs
```

## JACK
## Filtering out Jack from the pre-focals dataset
```{r}
Jack<- pre_focals %>% filter(FOCAL.ID == "JA")
```

### Getting SD for each month for Jack during pre-release observations
```{r}
#Month 6
Jack_behaviorcount_month6<- Jack %>% 
  filter(MONTH == "6") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month6

Jack_month6_SD<- diversity(Jack_behaviorcount_month6$count)
Jack_month6_SD

#Month 7
Jack_behaviorcount_month7<- Jack %>% 
  filter(MONTH == "7") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month7

Jack_month7_SD<- diversity(Jack_behaviorcount_month7$count)
Jack_month7_SD

#Month 8
Jack_behaviorcount_month8<- Jack %>% 
  filter(MONTH == "8") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month8

Jack_month8_SD<- diversity(Jack_behaviorcount_month8$count)
Jack_month8_SD

#Month 9
Jack_behaviorcount_month9<- Jack %>% 
  filter(MONTH == "9") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month9

Jack_month9_SD<- diversity(Jack_behaviorcount_month9$count)
Jack_month9_SD

#Month 10
Jack_behaviorcount_month10<- Jack %>% 
  filter(MONTH == "10") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month10

Jack_month10_SD<- diversity(Jack_behaviorcount_month10$count)
Jack_month10_SD

#Month 11
Jack_behaviorcount_month11<- Jack %>% 
  filter(MONTH == "11") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month11

Jack_month11_SD<- diversity(Jack_behaviorcount_month11$count)
Jack_month11_SD

#Month 12
Jack_behaviorcount_month12<- Jack %>% 
  filter(MONTH == "12") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month12

Jack_month12_SD<- diversity(Jack_behaviorcount_month12$count)
Jack_month12_SD

#Month 1
Jack_behaviorcount_month1<- Jack %>% 
  filter(MONTH == "1") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month1

Jack_month1_SD<- diversity(Jack_behaviorcount_month1$count)
Jack_month1_SD

#Month 2
Jack_behaviorcount_month2<- Jack %>% 
  filter(MONTH == "2") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Jack_behaviorcount_month2

Jack_month2_SD<- diversity(Jack_behaviorcount_month2$count)
Jack_month2_SD
```

### Averaging the SDs across Jack's 4 months pre-release
```{r}
Jack_pre_allSDs<- c(Jack_month6_SD,Jack_month7_SD,Jack_month8_SD,Jack_month9_SD,Jack_month10_SD,Jack_month11_SD,Jack_month12_SD,Jack_month1_SD,Jack_month2_SD)
Jack_pre_avgSDs<- mean(Jack_pre_allSDs)
#Jack_pre_avgSDs
```

## BLUE
## Filtering out Blue from the pre-focals dataset
```{r}
Blue<- pre_focals %>% filter(FOCAL.ID == "BL")
```

### Getting SD for each month for Blue during pre-release observations
```{r}
#Month 6
Blue_behaviorcount_month6<- Blue %>% 
  filter(MONTH == "6") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month6

Blue_month6_SD<- diversity(Blue_behaviorcount_month6$count)
Blue_month6_SD

#Month 7
Blue_behaviorcount_month7<- Blue %>% 
  filter(MONTH == "7") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month7

Blue_month7_SD<- diversity(Blue_behaviorcount_month7$count)
Blue_month7_SD

#Month 8
Blue_behaviorcount_month8<- Blue %>% 
  filter(MONTH == "8") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month8

Blue_month8_SD<- diversity(Blue_behaviorcount_month8$count)
Blue_month8_SD

#Month 9
Blue_behaviorcount_month9<- Blue %>% 
  filter(MONTH == "9") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month9

Blue_month9_SD<- diversity(Blue_behaviorcount_month9$count)
Blue_month9_SD

#Month 10
Blue_behaviorcount_month10<- Blue %>% 
  filter(MONTH == "10") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month10

Blue_month10_SD<- diversity(Blue_behaviorcount_month10$count)
Blue_month10_SD

#Month 11
Blue_behaviorcount_month11<- Blue %>% 
  filter(MONTH == "11") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month11

Blue_month11_SD<- diversity(Blue_behaviorcount_month11$count)
Blue_month11_SD

#Month 12
Blue_behaviorcount_month12<- Blue %>% 
  filter(MONTH == "12") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month12

Blue_month12_SD<- diversity(Blue_behaviorcount_month12$count)
Blue_month12_SD

#Month 1
Blue_behaviorcount_month1<- Blue %>% 
  filter(MONTH == "1") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month1

Blue_month1_SD<- diversity(Blue_behaviorcount_month1$count)
Blue_month1_SD

#Month 2
Blue_behaviorcount_month2<- Blue %>% 
  filter(MONTH == "2") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Blue_behaviorcount_month2

Blue_month2_SD<- diversity(Blue_behaviorcount_month2$count)
Blue_month2_SD
```

### Averaging the SDs across Blue's 4 months pre-release
```{r}
Blue_pre_allSDs<- c(Blue_month6_SD,Blue_month7_SD,Blue_month8_SD,Blue_month9_SD,Blue_month10_SD,Blue_month11_SD,Blue_month12_SD,Blue_month1_SD,Blue_month2_SD)
Blue_pre_avgSDs<- mean(Blue_pre_allSDs)
#Blue_pre_avgSDs
```

## ALEX
## Filtering out Alex from the pre-focals dataset
```{r}
Alex<- pre_focals %>% filter(FOCAL.ID == "AL")
```

### Getting SD for each month for Alex during pre-release observations
```{r}
#Month 6
Alex_behaviorcount_month6<- Alex %>% 
  filter(MONTH == "6") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month6

Alex_month6_SD<- diversity(Alex_behaviorcount_month6$count)
Alex_month6_SD

#Month 7
Alex_behaviorcount_month7<- Alex %>% 
  filter(MONTH == "7") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month7

Alex_month7_SD<- diversity(Alex_behaviorcount_month7$count)
Alex_month7_SD

#Month 8
Alex_behaviorcount_month8<- Alex %>% 
  filter(MONTH == "8") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month8

Alex_month8_SD<- diversity(Alex_behaviorcount_month8$count)
Alex_month8_SD

#Month 9
Alex_behaviorcount_month9<- Alex %>% 
  filter(MONTH == "9") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month9

Alex_month9_SD<- diversity(Alex_behaviorcount_month9$count)
Alex_month9_SD

#Month 10
Alex_behaviorcount_month10<- Alex %>% 
  filter(MONTH == "10") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month10

Alex_month10_SD<- diversity(Alex_behaviorcount_month10$count)
Alex_month10_SD

#Month 11
Alex_behaviorcount_month11<- Alex %>% 
  filter(MONTH == "11") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month11

Alex_month11_SD<- diversity(Alex_behaviorcount_month11$count)
Alex_month11_SD

#Month 12
Alex_behaviorcount_month12<- Alex %>% 
  filter(MONTH == "12") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month12

Alex_month12_SD<- diversity(Alex_behaviorcount_month12$count)
Alex_month12_SD

#Month 1
Alex_behaviorcount_month1<- Alex %>% 
  filter(MONTH == "1") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month1

Alex_month1_SD<- diversity(Alex_behaviorcount_month1$count)
Alex_month1_SD

#Month 2
Alex_behaviorcount_month2<- Alex %>% 
  filter(MONTH == "2") %>%
  gather("Obs", "behavior", 22:23, na.rm = T) %>%
  group_by(behavior) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
Alex_behaviorcount_month2

Alex_month2_SD<- diversity(Alex_behaviorcount_month2$count)
Alex_month2_SD
```

### Averaging the SDs across Alex's 4 months pre-release
```{r}
Alex_pre_allSDs<- c(Alex_month6_SD,Alex_month7_SD,Alex_month8_SD,Alex_month9_SD,Alex_month10_SD,Alex_month11_SD,Alex_month12_SD,Alex_month1_SD,Alex_month2_SD)
Alex_pre_avgSDs<- mean(Alex_pre_allSDs)
#Alex_pre_avgSDs
```

## BART
### Getting SD for each month for Bart during pre-release observations
```{r}
#Month 6
Bart_month6_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "6")
Bart_month6_SD<- diversity(Bart_month6_SD$count)

#Month 7
Bart_month7_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "7")
Bart_month7_SD<- diversity(Bart_month7_SD$count)

#Month 8
Bart_month8_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "8")
Bart_month8_SD<- diversity(Bart_month8_SD$count)

#Month 9
Bart_month9_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "9")
Bart_month9_SD<- diversity(Bart_month9_SD$count)

#Month 10
Bart_month10_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "10")
Bart_month10_SD<- diversity(Bart_month10_SD$count)

#Month 11
Bart_month11_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "11")
Bart_month11_SD<- diversity(Bart_month11_SD$count)

#Month 12
Bart_month12_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "12")
Bart_month12_SD<- diversity(Bart_month12_SD$count)

#Month 1
Bart_month1_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "1")
Bart_month1_SD<- diversity(Bart_month1_SD$count)

#Month 2
Bart_month2_SD<- monkey.prelim_SD2$BA %>%
  filter(MONTH == "2")
Bart_month2_SD<- diversity(Bart_month2_SD$count)
```

### Averaging the SDs across Bart's 4 months pre-release
```{r}
Bart_pre_allSDs<- c(Bart_month6_SD,Bart_month7_SD,Bart_month8_SD,Bart_month9_SD,Bart_month10_SD,Bart_month11_SD,Bart_month12_SD,Bart_month1_SD,Bart_month2_SD)
Bart_pre_avgSDs<- mean(Bart_pre_allSDs)
Bart_pre_avgSDs #1.38
```

## Neville
### Getting SD for each month for Neville during pre-release observations
```{r}
#Month 6
Nev_month6_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "6")
Nev_month6_SD<- diversity(Nev_month6_SD$count)

#Month 7
Nev_month7_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "7")
Nev_month7_SD<- diversity(Nev_month7_SD$count)

#Month 8
Nev_month8_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "8")
Nev_month8_SD<- diversity(Nev_month8_SD$count)

#Month 9
Nev_month9_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "9")
Nev_month9_SD<- diversity(Nev_month9_SD$count)

#Month 10
Nev_month10_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "10")
Nev_month10_SD<- diversity(Nev_month10_SD$count)

#Month 11
Nev_month11_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "11")
Nev_month11_SD<- diversity(Nev_month11_SD$count)

#Month 12
Nev_month12_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "12")
Nev_month12_SD<- diversity(Nev_month12_SD$count)

#Month 1
Nev_month1_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "1")
Nev_month1_SD<- diversity(Nev_month1_SD$count)

#Month 2
Nev_month2_SD<- monkey.prelim_SD2$NE %>%
  filter(MONTH == "2")
Nev_month2_SD<- diversity(Nev_month2_SD$count)
```

### Averaging the SDs across Nev's 4 months pre-release
```{r}
Nev_pre_allSDs<- c(Nev_month6_SD,Nev_month7_SD,Nev_month8_SD,Nev_month9_SD,Nev_month10_SD,Nev_month11_SD,Nev_month12_SD,Nev_month1_SD,Nev_month2_SD)
Nev_pre_avgSDs<- mean(Nev_pre_allSDs)
Nev_pre_avgSDs
```

## AUGUSTINE
### Getting SD for each month for Aug during pre-release observations
```{r}
#Month 6
Aug_month6_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "6")
Aug_month6_SD<- diversity(Aug_month6_SD$count)

#Month 7
Aug_month7_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "7")
Aug_month7_SD<- diversity(Aug_month7_SD$count)

#Month 8
Aug_month8_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "8")
Aug_month8_SD<- diversity(Aug_month8_SD$count)

#Month 9
Aug_month9_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "9")
Aug_month9_SD<- diversity(Aug_month9_SD$count)

#Month 10
Aug_month10_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "10")
Aug_month10_SD<- diversity(Aug_month10_SD$count)

#Month 11
Aug_month11_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "11")
Aug_month11_SD<- diversity(Aug_month11_SD$count)

#Month 12
Aug_month12_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "12")
Aug_month12_SD<- diversity(Aug_month12_SD$count)

#Month 1
Aug_month1_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "1")
Aug_month1_SD<- diversity(Aug_month1_SD$count)

#Month 2
Aug_month2_SD<- monkey.prelim_SD2$AU %>%
  filter(MONTH == "2")
Aug_month2_SD<- diversity(Aug_month2_SD$count)
```

### Averaging the SDs across Aug's 4 months pre-release
```{r}
Aug_pre_allSDs<- c(Aug_month6_SD,Aug_month7_SD,Aug_month8_SD,Aug_month9_SD,Aug_month10_SD,Aug_month11_SD,Aug_month12_SD,Aug_month1_SD,Aug_month2_SD)
Aug_pre_avgSDs<- mean(Aug_pre_allSDs)
Aug_pre_avgSDs 
```

## AMY
### Getting SD for each month for Amy during pre-release observations
```{r}
#Month 6
Amy_month6_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "6")
Amy_month6_SD<- diversity(Amy_month6_SD$count)

#Month 7
Amy_month7_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "7")
Amy_month7_SD<- diversity(Amy_month7_SD$count)

#Month 8
Amy_month8_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "8")
Amy_month8_SD<- diversity(Amy_month8_SD$count)

#Month 9
Amy_month9_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "9")
Amy_month9_SD<- diversity(Amy_month9_SD$count)

#Month 10
Amy_month10_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "10")
Amy_month10_SD<- diversity(Amy_month10_SD$count)

#Month 11
Amy_month11_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "11")
Amy_month11_SD<- diversity(Amy_month11_SD$count)

#Month 12
Amy_month12_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "12")
Amy_month12_SD<- diversity(Amy_month12_SD$count)

#Month 1
Amy_month1_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "1")
Amy_month1_SD<- diversity(Amy_month1_SD$count)

#Month 2
Amy_month2_SD<- monkey.prelim_SD2$AM %>%
  filter(MONTH == "2")
Amy_month2_SD<- diversity(Amy_month2_SD$count)
```

### Averaging the SDs across Amy's 4 months pre-release
```{r}
Amy_pre_allSDs<- c(Amy_month6_SD,Amy_month7_SD,Amy_month8_SD,Amy_month9_SD,Amy_month10_SD,Amy_month12_SD,Amy_month11_SD,Amy_month1_SD,Amy_month2_SD)
Amy_pre_avgSDs<- mean(Amy_pre_allSDs)
Amy_pre_avgSDs
```

## MAY
### Getting SD for each month for May during pre-release observations
```{r}
#Month 6
May_month6_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "6")
May_month6_SD<- diversity(May_month6_SD$count)

#Month 7
May_month7_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "7")
May_month7_SD<- diversity(May_month7_SD$count)

#Month 8
May_month8_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "8")
May_month8_SD<- diversity(May_month8_SD$count)

#Month 9
May_month9_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "9")
May_month9_SD<- diversity(May_month9_SD$count)

#Month 10
May_month10_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "10")
May_month10_SD<- diversity(May_month10_SD$count)

#Month 11
May_month11_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "11")
May_month11_SD<- diversity(May_month11_SD$count)

#Month 12
May_month12_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "12")
May_month12_SD<- diversity(May_month12_SD$count)

#Month 1
May_month1_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "1")
May_month1_SD<- diversity(May_month1_SD$count)

#Month 2
May_month2_SD<- monkey.prelim_SD2$MA %>%
  filter(MONTH == "2")
May_month2_SD<- diversity(May_month2_SD$count)
```

### Averaging the SDs across May's 4 months pre-release
```{r}
May_pre_allSDs<- c(May_month6_SD,May_month7_SD,May_month8_SD,May_month9_SD,May_month10_SD,May_month11_SD,May_month12_SD,May_month1_SD,May_month2_SD)
May_pre_avgSDs<- mean(May_pre_allSDs)
May_pre_avgSDs
```

## TONI
### Getting SD for each month for Toni during pre-release observations
```{r}
#Month 6
Toni_month6_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "6")
Toni_month6_SD<- diversity(Toni_month6_SD$count)

#Month 7
Toni_month7_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "7")
Toni_month7_SD<- diversity(Toni_month7_SD$count)

#Month 8
Toni_month8_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "8")
Toni_month8_SD<- diversity(Toni_month8_SD$count)

#Month 9
Toni_month9_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "9")
Toni_month9_SD<- diversity(Toni_month9_SD$count)

#Month 10
Toni_month10_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "10")
Toni_month10_SD<- diversity(Toni_month10_SD$count)

#Month 11
Toni_month11_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "11")
Toni_month11_SD<- diversity(Toni_month11_SD$count)

#Month 12
Toni_month12_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "12")
Toni_month12_SD<- diversity(Toni_month12_SD$count)

#Month 1
Toni_month1_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "1")
Toni_month1_SD<- diversity(Toni_month1_SD$count)

#Month 2
Toni_month2_SD<- monkey.prelim_SD2$TO %>%
  filter(MONTH == "2")
Toni_month2_SD<- diversity(Toni_month2_SD$count)
```

### Averaging the SDs across Toni's 4 months pre-release
```{r}
Toni_pre_allSDs<- c(Toni_month6_SD,Toni_month7_SD,Toni_month8_SD,Toni_month9_SD,Toni_month10_SD,Toni_month11_SD,Toni_month12_SD,Toni_month1_SD,Toni_month2_SD)
Toni_pre_avgSDs<- mean(Toni_pre_allSDs)
Toni_pre_avgSDs
```

## BOO
### Getting SD for each month for Boo during pre-release observations
```{r}
#Month 6
Boo_month6_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "6")
Boo_month6_SD<- diversity(Boo_month6_SD$count)

#Month 7
Boo_month7_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "7")
Boo_month7_SD<- diversity(Boo_month7_SD$count)

#Month 8
Boo_month8_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "8")
Boo_month8_SD<- diversity(Boo_month8_SD$count)

#Month 9
Boo_month9_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "9")
Boo_month9_SD<- diversity(Boo_month9_SD$count)

#Month 10
Boo_month10_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "10")
Boo_month10_SD<- diversity(Boo_month10_SD$count)

#Month 11
Boo_month11_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "11")
Boo_month11_SD<- diversity(Boo_month11_SD$count)

#Month 12
Boo_month12_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "12")
Boo_month12_SD<- diversity(Boo_month12_SD$count)

#Month 1
Boo_month1_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "1")
Boo_month1_SD<- diversity(Boo_month1_SD$count)

#Month 2
Boo_month2_SD<- monkey.prelim_SD2$BO %>%
  filter(MONTH == "2")
Boo_month2_SD<- diversity(Boo_month2_SD$count)
```

### Averaging the SDs across Boo's 4 months pre-release
```{r}
Boo_pre_allSDs<- c(Boo_month6_SD,Boo_month7_SD,Boo_month8_SD,Boo_month9_SD,Boo_month10_SD,Boo_month11_SD,Boo_month12_SD,Boo_month1_SD,Boo_month2_SD)
Boo_pre_avgSDs<- mean(Boo_pre_allSDs)
Boo_pre_avgSDs
```

## BIG MAMA
### Getting SD for each month for Bgm during pre-release observations
```{r}
#Month 6
Bgm_month6_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "6")
Bgm_month6_SD<- diversity(Bgm_month6_SD$count)

#Month 7
Bgm_month7_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "7")
Bgm_month7_SD<- diversity(Bgm_month7_SD$count)

#Month 8
Bgm_month8_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "8")
Bgm_month8_SD<- diversity(Bgm_month8_SD$count)

#Month 9
Bgm_month9_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "9")
Bgm_month9_SD<- diversity(Bgm_month9_SD$count)

#Month 10
Bgm_month10_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "10")
Bgm_month10_SD<- diversity(Bgm_month10_SD$count)

#Month 11
Bgm_month11_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "11")
Bgm_month11_SD<- diversity(Bgm_month11_SD$count)

#Month 12
Bgm_month12_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "12")
Bgm_month12_SD<- diversity(Bgm_month12_SD$count)

#Month 1
Bgm_month1_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "1")
Bgm_month1_SD<- diversity(Bgm_month1_SD$count)

#Month 2
Bgm_month2_SD<- monkey.prelim_SD2$BM %>%
  filter(MONTH == "2")
Bgm_month2_SD<- diversity(Bgm_month2_SD$count)
```

### Averaging the SDs across Bgm's 4 months pre-release
```{r}
Bgm_pre_allSDs<- c(Bgm_month6_SD,Bgm_month7_SD,Bgm_month8_SD,Bgm_month9_SD,Bgm_month10_SD,Bgm_month11_SD,Bgm_month12_SD,Bgm_month1_SD,Bgm_month2_SD)
Bgm_pre_avgSDs<- mean(Bgm_pre_allSDs)
Bgm_pre_avgSDs
```

## KOVU
### Getting SD for each month for Kovu during pre-release observations
```{r}
#Month 6
Kovu_month6_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "6")
Kovu_month6_SD<- diversity(Kovu_month6_SD$count)

#Month 7
Kovu_month7_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "7")
Kovu_month7_SD<- diversity(Kovu_month7_SD$count)

#Month 8
Kovu_month8_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "8")
Kovu_month8_SD<- diversity(Kovu_month8_SD$count)

#Month 9
Kovu_month9_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "9")
Kovu_month9_SD<- diversity(Kovu_month9_SD$count)

#Month 10
Kovu_month10_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "10")
Kovu_month10_SD<- diversity(Kovu_month10_SD$count)

#Month 11
Kovu_month11_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "11")
Kovu_month11_SD<- diversity(Kovu_month11_SD$count)

#Month 12
Kovu_month12_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "12")
Kovu_month12_SD<- diversity(Kovu_month12_SD$count)

#Month 1
Kovu_month1_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "1")
Kovu_month1_SD<- diversity(Kovu_month1_SD$count)

#Month 2
Kovu_month2_SD<- monkey.prelim_SD2$KO %>%
  filter(MONTH == "2")
Kovu_month2_SD<- diversity(Kovu_month2_SD$count)
```

### Averaging the SDs across Kovu's 4 months pre-release
```{r}
Kovu_pre_allSDs<- c(Kovu_month6_SD,Kovu_month7_SD,Kovu_month8_SD,Kovu_month9_SD,Kovu_month10_SD,Kovu_month11_SD,Kovu_month12_SD,Kovu_month1_SD,Kovu_month2_SD)
Kovu_pre_avgSDs<- mean(Kovu_pre_allSDs)
Kovu_pre_avgSDs
```

## EDDY
### Getting SD for each month for Eddy during pre-release observations
```{r}
#Month 6
Ed_month6_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "6")
Ed_month6_SD<- diversity(Ed_month6_SD$count)

#Month 7
Ed_month7_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "7")
Ed_month7_SD<- diversity(Ed_month7_SD$count)

#Month 8
Ed_month8_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "8")
Ed_month8_SD<- diversity(Ed_month8_SD$count)

#Month 9
Ed_month9_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "9")
Ed_month9_SD<- diversity(Ed_month9_SD$count)

#Month 10
Ed_month10_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "10")
Ed_month10_SD<- diversity(Ed_month10_SD$count)

#Month 11
Ed_month11_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "11")
Ed_month11_SD<- diversity(Ed_month11_SD$count)

#Month 12
Ed_month12_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "12")
Ed_month12_SD<- diversity(Ed_month12_SD$count)

#Month 1
Ed_month1_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "1")
Ed_month1_SD<- diversity(Ed_month1_SD$count)

#Month 2
Ed_month2_SD<- monkey.prelim_SD2$ED %>%
  filter(MONTH == "2")
Ed_month2_SD<- diversity(Ed_month2_SD$count)
```

### Averaging the SDs across Eddy's 4 months pre-release
```{r}
Ed_pre_allSDs<- c(Ed_month6_SD,Ed_month7_SD,Ed_month8_SD,Ed_month9_SD,Ed_month10_SD,Ed_month11_SD,Ed_month12_SD,Ed_month1_SD,Ed_month2_SD)
Ed_pre_avgSDs<- mean(Ed_pre_allSDs)
Ed_pre_avgSDs
```

## TINKER
### Getting SD for each month for Tinker during pre-release observations
```{r}
#Month 6
Tink_month6_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "6")
Tink_month6_SD<- diversity(Tink_month6_SD$count)

#Month 7
Tink_month7_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "7")
Tink_month7_SD<- diversity(Tink_month7_SD$count)

#Month 8
Tink_month8_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "8")
Tink_month8_SD<- diversity(Tink_month8_SD$count)

#Month 9
Tink_month9_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "9")
Tink_month9_SD<- diversity(Tink_month9_SD$count)

#Month 10
Tink_month10_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "10")
Tink_month10_SD<- diversity(Tink_month10_SD$count)

#Month 11
Tink_month11_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "11")
Tink_month11_SD<- diversity(Tink_month11_SD$count)

#Month 12
Tink_month12_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "12")
Tink_month12_SD<- diversity(Tink_month12_SD$count)

#Month 1
Tink_month1_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "1")
Tink_month1_SD<- diversity(Tink_month1_SD$count)

#Month 2
Tink_month2_SD<- monkey.prelim_SD2$TI %>%
  filter(MONTH == "2")
Tink_month2_SD<- diversity(Tink_month2_SD$count)
```

### Averaging the SDs across Tinker's 4 months pre-release
```{r}
Tink_pre_allSDs<- c(Tink_month6_SD,Tink_month7_SD,Tink_month8_SD,Tink_month9_SD,Tink_month10_SD,Tink_month11_SD,Tink_month12_SD,Tink_month1_SD,Tink_month2_SD)
Tink_pre_avgSDs<- mean(Tink_pre_allSDs)
Tink_pre_avgSDs
```

## BATMAN
### Getting SD for each month for Batman during pre-release observations
```{r}
#Month 6
Bat_month6_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "6")
Bat_month6_SD<- diversity(Bat_month6_SD$count)

#Month 7
Bat_month7_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "7")
Bat_month7_SD<- diversity(Bat_month7_SD$count)

#Month 8
Bat_month8_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "8")
Bat_month8_SD<- diversity(Bat_month8_SD$count)

#Month 9
Bat_month9_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "9")
Bat_month9_SD<- diversity(Bat_month9_SD$count)

#Month 10
Bat_month10_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "10")
Bat_month10_SD<- diversity(Bat_month10_SD$count)

#Month 11
Bat_month11_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "11")
Bat_month11_SD<- diversity(Bat_month11_SD$count)

#Month 12
Bat_month12_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "12")
Bat_month12_SD<- diversity(Bat_month12_SD$count)

#Month 1
Bat_month1_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "1")
Bat_month1_SD<- diversity(Bat_month1_SD$count)

#Month 2
Bat_month2_SD<- monkey.prelim_SD2$BT %>%
  filter(MONTH == "2")
Bat_month2_SD<- diversity(Bat_month2_SD$count)
```

### Averaging the SDs across Batman's 4 months pre-release
```{r}
Bat_pre_allSDs<- c(Bat_month6_SD,Bat_month7_SD,Bat_month8_SD,Bat_month9_SD,Bat_month10_SD,Bat_month11_SD,Bat_month12_SD,Bat_month1_SD,Bat_month2_SD)
Bat_pre_avgSDs<- mean(Bat_pre_allSDs)
Bat_pre_avgSDs
```

## MANGO
### Getting SD for each month for Mango during pre-release observations
```{r}
#Month 6
Mango_month6_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "6")
Mango_month6_SD<- diversity(Mango_month6_SD$count)

#Month 7
Mango_month7_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "7")
Mango_month7_SD<- diversity(Mango_month7_SD$count)

#Month 8
Mango_month8_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "8")
Mango_month8_SD<- diversity(Mango_month8_SD$count)

#Month 9
Mango_month9_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "9")
Mango_month9_SD<- diversity(Mango_month9_SD$count)

#Month 10
Mango_month10_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "10")
Mango_month10_SD<- diversity(Mango_month10_SD$count)

#Month 11
Mango_month11_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "11")
Mango_month11_SD<- diversity(Mango_month11_SD$count)

#Month 12
Mango_month12_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "12")
Mango_month12_SD<- diversity(Mango_month12_SD$count)

#Month 1
Mango_month1_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "1")
Mango_month1_SD<- diversity(Mango_month1_SD$count)

#Month 2
Mango_month2_SD<- monkey.prelim_SD2$MG %>%
  filter(MONTH == "2")
Mango_month2_SD<- diversity(Mango_month2_SD$count)
```

### Averaging the SDs across Mango's 4 months pre-release
```{r}
Mango_pre_allSDs<- c(Mango_month6_SD,Mango_month7_SD,Mango_month8_SD,Mango_month9_SD,Mango_month10_SD,Mango_month11_SD,Mango_month12_SD,Mango_month1_SD,Mango_month2_SD)
Mango_pre_avgSDs<- mean(Mango_pre_allSDs)
Mango_pre_avgSDs
```

### Full Troop SD DF and Plot
```{r}
#Creating a matrix
pre_SDs_table<- matrix(c(Pop_pre_avgSDs, Zip_pre_avgSDs, Jack_pre_avgSDs, Blue_pre_avgSDs, Alex_pre_avgSDs, Bart_pre_avgSDs, Nev_pre_avgSDs, Aug_pre_avgSDs, Amy_pre_avgSDs, May_pre_avgSDs, Toni_pre_avgSDs, Boo_pre_avgSDs, Bgm_pre_avgSDs, Kovu_pre_avgSDs, Ed_pre_avgSDs, Tink_pre_avgSDs, Bat_pre_avgSDs, Mango_pre_avgSDs), ncol = 1, byrow = 18)
pre_SDs_table

#Turning into dataframe and adding new columns
pre_SDs_table<- as.data.frame(pre_SDs_table) %>%
  rename(SD.Index = V1) %>%
  add_column(IDs = c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Neville", "Augustine", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman", "Mango")) %>%
  add_column(release = "pre")

PR_SD_PLOT<- barplot(pre_SDs_table, beside = TRUE, space = c(0.25,1), xlab = "", ylab = "Shannon Diversity Index", names.arg = c("Pops","Zip","Jack", "Blue", "Alex", "Bart", "Nev", "Aug", "Amy", "May", "Toni", "Boo", "Big Mama", "Kovu", "Eddy", "Tinker", "Batman", "Mango"), ylim = c(0,2), cex.names = 0.8, las=2, col="#339999")#I don't love this barplot visual, I can't figure out how to turn the x-axis labels on an angle using a table
PR_SD_PLOT

#using ggplot
SD_plot<- ggplot(data= pre_SDs_table,aes(x=IDs, y=SD.Index)) +
  geom_bar(position="dodge", stat="identity") +
  theme(axis.text.x = element_text(angle = 45, hjust =1)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))
```


